---
title: "CPS Longitudinal - Analytic Plan"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-05-05"
bibliography: "refs.bib"
knit: |
  (function(input, ...) {
      rmarkdown::render(
        input, 
        output_file   = paste0("33753_sow3_analytic_plan_", Sys.Date()),
        output_dir    = file.path(dirname(input), "output"),
        envir         = globalenv()
      )
  })
---

look at 

- [ ] assns  between variables during same year
- [ ] lagged assn between variables during supsequent years





after https://www.andrewheiss.com/blog/2020/12/01/ipw-binary-continuous/

(cross-sectional) iptw: 

numer_mod = glm(util ~ 1, data = data)
numer = dnorm(data$util, predict(numer_nod), sd = numer_mod$residuals)

denom_mod = glm(util ~ demog + asthma, data = data)
denom = dnorm(data$util, predict(denom_nod), sd = denom_mod$residuals)

data$ipw = numer / denom
attendance_mod = glm(attendance ~ util, data = data, weights = ipw)

alt: use ipw::ipwpoint()





# Difference-in-Differences

@Athey2022




# Missing Data

@King2008 offers an example of using complete reference data to infer 
missnig categorical characteristics. Distributions of both 
outcome variables and indicator variables may differ in the observed (complete)
and inferred (incomplete) subpopulations. 

In the @King2008, treat the cause-of-death value of interest [P(D)] as 
multinomial. Estimate, e.g., probability of each 
symptom profile (fully observed var) given observed cause of death (partially observed var):

P(S) = P(S|D)P(D)

Assume P(S|D) are the same across subpopulations. Don't assume P(S), P(D),
or P(D|S) are the same. Tabulate observed values nonparametrically
and solve for P(D). Use samples of symptoms, solve, and average. [Applicable
to uninferrable xtabs?]

# Causal Inference




" Nonadherence is often related to social determinants of health, making it even more important to study nonadherence patterns and to estimate both the ITT effect and the per-protocol effect. " [@RojasSaunero2022]

"In fact, our recommendation is to anticipate as many of these adjacent questions as possible whenever designing a randomized trial of a social intervention, and to collect data so that their effect estimates can be explored and estimated (with additional assumptions) by emulating the adjacent target trials within our conducted trials’ data. Ultimately, the effort to collect rich data in a trial setting can serve as an opportunity to estimate the adapted interventions of treatment strategies beyond those that defined the original trial’s protocol. Note that such data collection requires forethought and at minimum needs to include assessments of the anticipated adapted intervention definitions and associated confounders (9, 15, 16). " [@RojasSaunero2022]

"Broadly speaking, interference means that the intervention in one individual may affect other individuals’ adherence and outcomes. [...] Interference implies there could be many questions of interest, such as direct effects of the intervention, spillover effects, and overall effects (17, 18)" [@RojasSaunero2022]


"No interference between units is a component of Rubin's Stable Unit Treatment Value Assumption or SUTVA (Rubin, 1980, 1986). The assumption will be violated in settings in which social interactions allow one individual's exposure or treatment assignment to affect the outcomes of other individuals. " [@Vanderweele2013]

"Sincewecanrarelyobservethesamepatientsunderbothconditions (eveninacrossover randomizedcontrolledtrial carryover effects induce systematic differences), the onlyway to determinecausaleffectsistocomparetwogroupsthatareidentical ineverywayexcept thatonegroupwas treatedandthe otherwasnot.Marginalstructuralmodels(asothercausalmodels) attempt to fully adjust for measured confounders to enhancegroupcomparabilityandestimatecausal effects ina similarway [8]" [@Williamson2017]

"Marginal structural models aim to appropriately control for the effects of time-varying confounders that are affected by prior treatment (exposure). Marginal structural models are called marginal because they use the marginal distribution of the treatment (or exposure) variable at any time point to weight its relationship with outcome. Marginal in statistics means ‘unconditional’—i.e. not conditional on other variables, like the probabilities of one variable on the margins of a contingency table of two discrete variables (see below). These models are called structural because models exploring causal relationships are referred to as structural in the econometric and social sciences literature [8]." [@Williamson2017]

"Of the different available methods that have been proposed to address time-varying confounding with marginal structural models [14], inverse-probability-of-treatment weighting (IPTW) is the most commonly used [15]. Marginal structural model estimation via IPTW occurs in two stages (Figure 5). As a preliminary step, researchers choose the time scale, and define time, event, treatment and other covariates, and their relationships (Figures 2 and 4). In the first stage IPTW are calculated; in the second stage the outcome model is fit, including sensitivity analyses that take into account weight distributions." [@Williamson2017]




[@Wal2011]




@Robins2000

"Many considerations influence the selection of one therapy over another. In many settings, more than one therapeutic approach is commonly used. In routine clinical practice, patients receiving one treatment will tend to be different from those receiving another, eg, if one treatment is thought to be better tolerated by elderly patients or more effective for patients who are more seriously ill. This results in a correlation—or confounding—between patient characteristics that affect outcomes and the choice of therapy (often called “confounding by indication”). If observational data obtained from routine clinical practice are examined to compare the outcomes of patients treated with different therapies, the observed difference will be the result of both differing patient characteristics and treatment choice, making it difficult to delineate the true effect of one treatment vs another." [@Haukoos2015]

--
Estimate propensity score by, e.g., logistic regression. Dependent variable
is probability that observational unit received treatment.

* Propensity score matching: compare units with similar propensity
scores but different treatment statuses; analyze as paired
* Stratification: Categorize based on propensity score and analyze within strata 
* Adjustment: 
* IPTW: 




" In the current study, we reverse the direction of the causal effect of interest 
and focus on the ATE of a non-randomized, latent, multinomial exposure 
(i.e., a latent class variable) on an observed, binary outcome. " [@Bray2018]

" Critically, if adolescent patterns of reasons for use are associated with later outcomes, even after using causal inference methods to account for differences across people with different patterns, it provides evidence that patterns of reasons could be used to screen individuals into prevention programs and/or tailor the content of programs. " [@Bray2018]




Class membership: school risk & utilization profile (e.g., high need, high use; high need, low use)
Outcome: 



what predicts uptake?
what predicts asthma?
does uptake predict asthma, after adjusting for predictors of uptake?




# Student Performance

What predicts utilization? Poisson (or similar) for inhaler uses, 
with a student-time (in program) offset. 
Adjust for school-level chars. 

- [ ] tabulate uses per school per year
- [ ] calculate student-program-time per school per year 
      (time btw program implementation and end of school year)
- [ ] calculate use rate per school per year
- [ ] plot 
    - [ ] scatter: rate over time by impl date
- [ ] model uses (count with student-time offset) 
      as predicted by school demographics,
      school characteristics, calendar time (school year), time in program
- [ ] consider LCA for racial composition? school chars?

Categorize schools by rate of utilization (quartiles). Categorize 
schools by outcome rates (quartiles). Tabulate, plot by year. 


Describe trends in student outcomes (student group level) 
over time (asthma, 504 plans, ieps, attendance).
Allow for demog-by-outcome interaction terms.


Latent class analysis/latent transition analysis. Categorize by low/high utilizers 
and low/high need. 

Describe trends in outcomes as predicted by time-in-program, utilization 
(treat as time-varying exposure?). Causal analysis w 
time-varying covariates (confounders): use IPTW for utilization?


```{r}
library(tidyverse)
library(lme4)
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}

```

```{r}
read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  summarise(
    n = sum(student_count),
    .by = c(school_id, data_schoolyear, student_race)
  ) |> 
  filter(!is.na(student_race)) |> 
  
  filter(school_id == first(school_id)) |> 
  pivot_wider(
    names_from = data_schoolyear,
    values_from = n, 
    id_cols = student_race,
    values_fill = 0
  ) |> 
  column_to_rownames("student_race") |> 
  as.matrix() |> 
  chisq.test()
```

```{r}
read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  pivot_longer(
    matches("student_504"),
    values_to = "n",
    names_to  = "y_level"
    ) |> 
  filter(n > 0) |> 
  summarise(
    n   = sum(n), 
    .by = c(
      school_id, data_schoolyear, 
      asthma_status, y_level, 
      student_race, student_gender
      )
    ) |> 
  mutate(
    p_obs        = case_when(!is.na(student_race) ~ n / sum(n)),
    n_missing    = sum(n*is.na(student_race)),
    n_nonmissing = sum(n*!is.na(student_race)),
    .by          = c(
      school_id, data_schoolyear, 
      asthma_status, y_level, 
      student_gender
      )
    ) |> 
  filter(!is.na(student_race))  |>
  arrange(
    school_id, data_schoolyear, 
    asthma_status, y_level, 
    student_gender, student_race
    ) |> 
  
  mutate(
    n_imp = c(
      rmultinom(
        n = 1,
        size = first(n_missing),
        prob = p_obs
      )),
    .by = c(school_id, data_schoolyear, asthma_status, y_level)
  ) |> 
  
  mutate(n = n + n_imp) #|> 
  pivot_wider(
    names_from = 
  )
  # 
  # pivot_wider(
  #   names_from = data_schoolyear,
  #   values_from = c(n_missing, p_nodem),
  #   # values_fill = 0,
  #   id_cols = c(school_id, asthma_status, y_level,  student_gender, student_race)
  #   ) |> 
  pivot_wider(
    names_from = student_race,
    values_from = p_obs,
    
    id_cols = c(school_id, asthma_status, y_level, data_schoolyear, n_nonmissing, n_missing)
  ) |> 
  arrange(school_id, asthma_status, y_level, data_schoolyear) 
  
  mutate(
    p_gender_g_row = n / sum(n), 
    .by = c(school_id, data_schoolyear,
            asthma_status, y_level,
            student_race)
    ) |> 
  mutate(
    p_race_g_row = n / sum(n), 
    .by = c(school_id, data_schoolyear,
            asthma_status, y_level,
            student_gender)
    )


```


```{r}
df1 <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  select(-student_count) |>  
  pivot_longer(
    matches("student_504"),
    values_to = "student_count",
    names_to  = "y_level"
    ) |> 
  pivot_wider(
    names_from = student_gender,
    values_from = student_count,
    values_fill = 0,
    id_cols = c(
      school_id, data_schoolyear, student_race,
      asthma_status, y_level
      )
  ) |> 
  janitor::clean_names() |> 
  mutate(
    n_obs = female + male + non_binary
    ) |> 
  filter(n_obs > 0) |> 
  mutate(across(where(negate(is.numeric)), as.factor))
```


```{r}
pred_m <- glmer(
  cbind(male, I(n_obs - male)) ~ 
    student_race + data_schoolyear + asthma_status + y_level + (1 | school_id),
  data   = droplevels(na.omit(df1)),
  family = binomial(link = "logit")
  )
```

```{r}
mod_gender_df <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |>
    select(-student_count) |>  
    pivot_longer(
      matches("student_504"),
      values_to = "student_count",
      names_to  = "y_level"
      ) |> 
    mutate(
      student_race = fct_na_value_to_level(student_race, "Missing")
      ) |> 
    filter(!is.na(student_gender))

library(mclogit)
mod_gender <- mclogit(
  student_count|student_gender ~ data_schoolyear + asthma_status + student_race + y_level,
  random = ~ 1 | school_id,
  data = group_nest(school_id)mod_gender_df
)
```


```{r}
pred_f <- glmer(
  cbind(female, I(n_total - female)) ~ 
        student_race + data_schoolyear + asthma_status + y_level + (1 | school_id),
  data   = droplevels(na.omit(df1)),
  family = binomial(link = "logit")
  )
```

```{r}
rmult_gender <- function(n_unobs, Female, Male, `Non-Binary`, ...) {
    data.frame(
      ...,
      student_count = rmultinom(
        n    = 1,
        size = n_unobs,
        prob = c("Female" = Female, "Male" = Male, "Non-Binary" = `Non-Binary`)
        )
      ) |> 
      rownames_to_column("student_gender")
}
```


```{r}
gender_missing <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  select(-student_count) |> 
  pivot_longer(
    matches("student_504"),
    values_to = "student_count",
    names_to  = "y_level"
    ) |> 
  pivot_wider(
    names_from  = student_gender,
    values_from = student_count,
    values_fill = 0,
    id_cols = c(
      school_id, data_schoolyear, student_race, asthma_status, y_level)
    ) |> 
  filter(!is.na(student_race))
gender_missing$pred_m <- predict(
  pred_m,
  newdata = gender_missing,
  type = "response"
  )
gender_missing$pred_f <- predict(
  pred_f,
  newdata = gender_missing,
  type = "response"
  )
```


```{r}
gender_imputed <- gender_missing |> 
  mutate(
    pred_total = pred_m + pred_f,
    pred_nb    = case_when(pred_total < 1 ~ (1 - pred_total), .default = 0.0000001)
    ) |> 
  janitor::clean_names() |> 
  # filter(na > 0) |>
  transpose() |> 
  map_dfr(~{
    n_addl <- c(rmultinom(
        n    = 1,
        size = .x[["na"]],
        prob = .x[c("pred_m", "pred_f", "pred_nb")]
        )
      ) |> set_names(c("male", "female", "non_binary"))
    data.frame(.x) |>
      mutate(
        Male         = male + n_addl[["male"]],
        Female       = female + n_addl[["female"]],
        `Non.Binary` = `non_binary` + n_addl[["non_binary"]]
      )

  })

gender_imputed |> 
  pivot_longer(
    c(Male, Female, Non.Binary), 
    values_to = "student_count",
    names_to = "student_gender"
    ) |> 
  pivot_wider(
    names_from = y_level,
    values_from = student_count,
    id_cols = c(school_id, data_schoolyear, student_race, asthma_status)
  )
  head()
```


```{r}
rmult_gender <- function(n_unobs, Female, Male, `Non-Binary`, ...) {
    data.frame(
      ...,
      student_count = rmultinom(
        n    = 1,
        size = n_unobs,
        prob = c("Female" = Female, "Male" = Male, "Non-Binary" = `Non-Binary`)
        )
      ) |> 
      rownames_to_column("student_gender")
}
```


```{r}
# for a single outcome, fill in missing covariate values

allocate_missing <- function(df, y = "student_504_count", x = "student_gender") {
  imputed <- df |> 
    select(-student_count) |> 
    pivot_longer(
      matches(y),
      values_to = "student_count",
      names_to  = "y_level"
      ) 
  
  imputed[["x_level"]] <- imputed[[x]]

  imputed <- imputed |> 
    mutate(
      n_obs   = sum(student_count*(!is.na(x_level))),
      n_unobs = sum(student_count*(is.na(x_level))),
      p_obs   = coalesce(student_count / (n_obs), 0),
      .by     = all_of(setdiff(
        c("school_id", "data_schoolyear", "student_race", "student_gender",
          "asthma_status", "y_level"),
        x))
      ) 
  return(imputed)
  #|> 
    filter(!is.na(x_level)) |> 
    filter(n_unobs > 0) |>
    # filter(p_obs > 0) |> 
    group_nest(
      across(all_of(setdiff(
        c("school_id", "data_schoolyear", "student_race", "student_gender",
          "asthma_status", "y_level", "n_unobs"
          ), x
        )))
      ) |> 
  mutate(
    n_rand = map2(n_unobs, data, possibly(~{
      data.frame(
        student_count = rmultinom(
          n    = 1,
          size = .x,
          prob = set_names(.y$p_obs, .y[[x]])
          )) |> 
        rownames_to_column(x)
      }, otherwise = data.frame(
        student_gender = "Undefined",
        student_count = -Inf
        )
      ))
    ) |> 
    unnest(n_rand) |> 
    mutate(
      student_count = case_when(
        student_gender == "Undefined" ~ n_unobs,
        .default = student_count
        )
      ) |> 
    pivot_wider(
      values_from = student_count,
      names_from = y_level,
      values_fill = 0,
      id_cols = c("school_id", "data_schoolyear", "student_race", "student_gender",
          "asthma_status", "n_unobs")
      
    )
  # return(imputed)
  full_join(
    df[which(!is.na(df[[x]])),intersect(names(df), names(imputed))],
    imputed
    ) |>
    arrange(school_id, data_schoolyear)
}
```


```{r}
read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  select(matches("student_504_count")) |> 
  colSums()
```


```{r}
list("iep_count", "student_504_count") |> 
  map(~{
    read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
      allocate_missing(y = .x)
    }) 

```


```{r}
#|> 
  filter(n_unobs > 0) |> 
  filter(p_obs > 0) |> 
  group_nest(
     school_id, data_schoolyear, student_race, asthma_status, y_level, n_unobs
     ) |> 
  mutate(
    n_rand = map2(n_unobs, data, ~{
      data.frame(
        student_count = rmultinom(
          n    = 1,
          size = .x,
          prob = set_names(.y$p_obs, .y[["student_gender"]])
          )) |> 
        rownames_to_column("student_gender")
      }), .keep = "unused") |> 
  unnest(n_rand)
```


```{r}
|> 
    pivot_wider(
      names_from  = x_level,
      values_from = p_obs,
      id_cols     = c(school_id, data_schoolyear, student_race, asthma_status, y_level,
                  n_gender_obs, n_gender_unobs),
      values_fill = 0
    ) |>
    filter(n_unobs > 0) |> 
    head() |> 
    pmap_dfr(allocate_gender) |> 
    
    pivot_wider(
      names_from  = y_level,
      values_from = student_count,
      id_cols     = c(school_id, data_schoolyear, student_race, asthma_status, student_gender)
    )
}
```


```{r}
df |> 
  select(school_id, data_schoolyear, student_gender, student_race, asthma_status,
         student_504_count_yes, student_504_count_no) |> 
  pivot_longer(
    matches("count"),
    values_to = "student_count",
    names_to  = "x504_status"
    ) |> 
  
  mutate(
    n_gender_obs   = sum(student_count*(!is.na(student_gender))),
    n_gender_unobs = sum(student_count*(is.na(student_gender))),
    p_gender       = student_count / n_gender_obs,
    .by            = c(school_id, data_schoolyear, student_race, asthma_status, x504_status)
  ) |> 
  filter(!is.na(student_gender)) |> 
  pivot_wider(
    names_from  = student_gender,
    values_from = p_gender,
    id_cols     = c(school_id, data_schoolyear, student_race, asthma_status, x504_status,
                n_gender_obs, n_gender_unobs),
    values_fill = 0
  ) |>
  filter(n_gender_unobs > 0) |> 
  head() |> 
  pmap_dfr(allocate_gender) |> 
  
  pivot_wider(
    names_from  = x504_status,
    values_from = student_count,
    id_cols     = c(school_id, data_schoolyear, student_race, asthma_status, student_gender)
  )

```













stratified by year:

limit to schools with complete data on sex.

for each outcome: pivot to long
```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```


P(S) = P(S|D)P(D)

P(race,asthma,504) = P(race,asthma,504|sex)*P(sex)

```{r}
# estimate p(vars) (fully observed)
head(p_vars <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  # filter(sum(is.na(student_gender))>0, .by = school_id) |> 
  # filter(as.numeric(factor(school_id)) == 1) |> 
  mutate(n_total = sum(student_count), .by = c(school_id, data_schoolyear)) |> 
  pivot_longer(
    matches("student_504_count"), 
    names_to  = "y_level", 
    values_to = "n"
    ) |> 
  summarise(
    n_row = sum(n), 
    .by   = c(
      school_id, data_schoolyear, n_total,
      student_race, asthma_status, y_level
      )
    ) |> 
  mutate(
    p_row = n_row / sum(n_row), 
    .by   = c(school_id, data_schoolyear, n_total)
  ))
```

```{r}

library(mclogit)

mblogit(formula = Opponent2 ~ Ethnicity2 + Rank2 + Opp_rank2, random = list(~1|Athlete2, ~ 1|Date2), data = data, method=c("PQL"), estimator=c("ML"))
```

```{r}
p_vars |> slice(100) |> left_join(p_vars_g_sex, by = c("school_id", "data_schoolyear", "student_race", "asthma_status", "y_level"))
```





```{r}

# for each sex level...
p_vars_g_sex <- levels(read_rds(nested_here("data", "analytic_student_lvl.rds")) |> pull("student_gender")) |> 
  set_names() |> 
  map_dfr(~{
    # estimate p(vars|sex) using observed data only
    read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
      filter(sum(is.na(student_gender))>0, .by = c(school_id, data_schoolyear)) |>
      mutate(
        gender_bin = case_when(student_gender == .x ~ "sex", .default = "not_sex")
      ) |> 
      # filter(as.numeric(factor(school_id)) == 1) |> 
      pivot_longer(
        matches("student_504_count"), 
        names_to  = "y_level", 
        values_to = "n"
        ) |> 
      summarise(
        n_row_if = sum(n), 
        .by = c(
          school_id, data_schoolyear, gender_bin,
          student_race, asthma_status, y_level
          )
        ) |> 
      mutate(
        p_row_if = n_row_if / sum(n_row_if), 
        .by = c(school_id, data_schoolyear, gender_bin)
        )
  }, .id = "student_gender")

```


P(race,asthma,504) = P(race,asthma,504|sex)*P(sex)



```{r}
p_vars_g_sex|> 
  pivot_wider(
    names_from  = gender_bin,
    values_from = c(n_row_if, p_row_if),
    values_fill = 0
  ) |> 
  left_join(
    p_vars, 
    by = c("school_id", "data_schoolyear", "student_race", "asthma_status", "y_level")
    ) |> 
  mutate(
    p_sex = (p_row - p_row_if_not_sex) / 
            (p_row_if_sex - p_row_if_not_sex),
    p_sex_g_row = p_sex / p_row
      
    )

```



```{r}
df <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  pivot_longer(
    matches("student_504_count"), 
    names_to  = "y_level", 
    values_to = "n"
    ) |> 
  # filter(!is.na(student_gender)) |> 
  summarise(
    n_row_if_sex = sum(n), .by = c(
      school_id, data_schoolyear, student_gender,
      student_race, asthma_status, y_level
      )
    ) |> 
  mutate(
    n_row_total = sum(n_row_if_sex*(!is.na(student_gender))),
    .by = c(
      school_id, data_schoolyear, 
      student_race, asthma_status, y_level
      )
    ) |> 
  
  filter(n_row_total > 0) |> 
  # filter(!grepl("Middle", student_race)) |> 
  droplevels()
library(lme4)
pred1 <- glmer(
  n_row_if_sex ~ student_race + data_schoolyear + (
    student_gender + asthma_status + y_level
    )^2 + (1 | school_id),
  data   = na.omit(df),
  family = poisson(link = "log"),
  offset = log(n_row_total)
)
```









```{r}
head(imp_test <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  filter(as.numeric(data_schoolyear) == 1) |> 
  droplevels())
```




```{r}
df$pred1 <- predict(pred1, newdata = df, type = "response")

df |> 
  ggplot(aes(n_row_if_sex, pred1)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0)


df |>
  filter(sum(is.na(student_gender)) > 0, .by = c(school_id, data_schoolyear)) |> 
  relocate(n_row_if_sex, n_row_total, pred1) |> 
  mutate(p_actual = n_row_if_sex / sum(n_row_if_sex),
         p_pred1 = pred1 / sum(pred1, na.rm = TRUE),
         .by = c(school_id, data_schoolyear, 
         student_race, asthma_status, y_level),
         .after = pred1
         ) |> 
  
  ggplot(aes(p_actual, p_pred1)) +
  geom_point()
```



P(race,asthma,504) = P(race,asthma,504|sex)*P(sex)
p(sex_i) = P(race,asthma,504) / P(race,asthma,504|sex_i)

```{r}
full_join(
  p_vars,
  p_vars_g_sex,
  by = c("school_id", "data_schoolyear", "student_race", "asthma_status",
"y_level")
) |> 
  mutate(
    n_row_if_notsex = n_row - n_row_if_sex,
    p_row_if_notsex = n_row_if_notsex / sum(n_row_if_notsex),
    .by = c(school_id, data_schoolyear, I(student_gender == student_gender))
    )
  select(-n_row_if_sex) |> 
  pivot_wider(
    names_from   = student_gender,
    values_from  = p_row_if_sex,
    names_prefix = "p_row_if_",
    values_fill  = 0
  ) |> 
  
  mutate(
    p_female = (p_row - p_row_if_Male)/ (p_row_if_Female - p_row_if_Male),
    p_male   = (p_row - p_row_if_Female)/ (p_row_if_Male - p_row_if_Female),
    p_total  = (p_female + p_male)
    )
  
  
```


```{r}
p_vars_g_sex |> 
  pivot_wider(
    names_from = student_gender,
    names_prefix = "p_vars_gvn_",
    values_from = p,
    id_cols = c(school_id, data_schoolyear,
                student_race, asthma_status, y_level),
    values_fill = 0
    ) |> 
  
  full_join(
    p_vars,
    by = c("school_id", "data_schoolyear", "student_race", "asthma_status",
"y_level")
  )

```

