---
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
output: word_document
title: "SOW 3 Analysis: Longitudinal CPS Data"
params:
  version: "2026-02-18"
  project_dir: 'a11_sow3_longitudinal' # relative to project root
  raw_data_cps:
    value:
      file: "UIC Asthma Status Data Albuterol SoW Shell_07252025.xlsx"
      sheet: "SY22, 23, 24 and 25 Asthma Stat"
      box_id: '1938297974039'
  raw_data_schools_metadata:
    value: 
      file: "CPS Schools CCAs Regions.xlsx"
      sheet: NULL
      box_id: '2071252778107'
  raw_data_utilization:
    value:
      file: "CPS 2023-2025 utilization recode.xlsx"
      sheet: "Sheet1"
      box_id: '2017350498669'
  raw_data_waves: # added 2026-01-21
    value:
      file: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
      box_id: "2109640911794"
  force_preproc: FALSE
  analytic_files:
    value:
      school_ids: "cps_ids.rds"
      student_counts: "student_counts.rds"
      school_chars: "school_chars.rds"
      school_year_dates: "school_year_dates.rds"
      inhaler_events: "inhaler_use_analytic.rds"
      inhaler_event_counts: "inhaler_use_counts.rds"
      school_counts_annual: "school_chars_annual.rds"
knit: |
  (function(input, ...) {
      rmarkdown::render(
        input, 
        output_format = rmarkdown::word_document(), 
        output_file   = paste0("33753_cps_sow_3_", Sys.Date(), ".docx"),
        output_dir    = file.path(dirname(input), "output"),
        envir         = globalenv()
      )
  })
---

```{r, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, eval = TRUE)
```

```{r}
# load libraries
library(tidyverse)
library(arsenal)
```

```{r}
## Set universal meta values
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# functions to simplify reading/writing relative paths
find_dated_code <- function(
    f_name,
    f_dir = here::here(params$project_dir, params$version, "R"),
    ext   = ".R"
    ) {
  f_name <- tools::file_path_sans_ext(f_name)
  file.path(f_dir, paste0(f_name, ext))
}

find_dated_output <- function(
    f_name, 
    f_dir = here::here(params$project_dir, params$version, "output"),
    ext   = ".rds"
    ) {
  f_name <- tools::file_path_sans_ext(f_name)
  file.path(f_dir, paste0(f_name, ext))
}

find_raw_data <- function(
    f_name,
    f_dir  = here::here(params$project_dir, "data-raw"),
    ...
    ) {
  if(f_name == "all") {
    f_name <- list.files(f_dir, ...)
    }
  file.path(f_dir, f_name)
}

find_analytic_data <- function(
    f_name, 
    f_dir = here::here(params$project_dir, params$version, "data"),
    ext   = ".rds"
    ) {
  f_name <- tools::file_path_sans_ext(f_name)
  file.path(f_dir, paste0(f_name, ext))
}

read_analytic <- function(
    f_name,
    f_dir = here::here(params$project_dir, params$version, "data"),
    ext   = ".rds"
    ) {
  f_name <- find_analytic_data(f_name, f_dir = f_dir, ext = ext)
  if(!file.exists(f_name)) {return(NULL)}
  res <- read_rds(f_name)
  res
}
```

```{r}
# create locations for output
here::here(params$project_dir, params$version, c("data", "output")) |> 
  walk(~{
    if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
    })
```

# Data Processing

```{r}
# if any analytic dataset doesn't exist, run preprocessing code
if (
  (any(!file.exists(find_analytic_data(params$analytic_files)))) |
    (params$force_preproc)
  ) {
  source(find_dated_code("r01_preproc"))
  }
```

- [x] Read data from Box. Raw files:

```{r}
params |> 
  keep_at(~ grepl("raw_data", .x)) |> 
  map_dfr(as.list, .id = "contents") |> 
  knitr::kable()
```

- [x] Match school names to school IDs in each dataset
* [x] incorporate updated implementation dates for schools (from Paige 2026-01-21)

<!-- * [ ] incorporate fxn to update asof dates automatically -->


* check data
   * [ ] missing or NA values
   * [ ] recodes

# Descriptive

* [ ] summarize school-level chars by school year:
    * [ ] n schools
    * [ ] wave
    * [ ] grade level
    * [ ] n students
    * [ ] region (all groups)
    * [ ] region (3 groups)
    * [ ] % students in each race and ethnicity category
    * [ ] % students in each gender category
    * [ ] % students ESL
    * [ ] % students FRM
    * [ ] % students homeless
    * [ ] average attendance
    * [ ] % students asthma
    * [ ] % students w food allergy
    * [ ] % students w non-food allergy
    * [ ] % students with 504
    * [ ] % students with IEP

```{r}

```




```{r}
# load cleaned datasets
# analytic_data <- params$analytic_files |> 
#   map(read_analytic)
```



```{r}
# extract utilization by school & year
# use_counts <- use_df |> 
  # summarise(
  #   n_events_sy = n(), 
  #   .by         = c(school_id, data_schoolyear)
  # ) |> 
  # right_join(
  #   school_df |> 
  #     distinct(school_id) |> 
  #     merge(
  #       data.frame(data_schoolyear = factor(c("2021-2022", "2022-2023", "2023-2024", 
  #                                             "2024-2025", "2025-2026"))),
  #       by  = NULL
  #     ), by = c("school_id", "data_schoolyear")
  #   ) |> 
  # 
  # mutate(n_events_sy = coalesce(n_events_sy, 0)) |> 
  # arrange(school_id, data_schoolyear) |> 
  # mutate(
  #   n_events_cum   = cumsum(n_events_sy),
  #   .by            = school_id
  # ) |> 
  # mutate(
  #   any_events_sy  = factor(
  #     n_events_sy > 0, 
  #     levels = c(FALSE, TRUE), 
  #     labels = c("0 Events", "1+ Events")
  #     ),
  #   any_events_cum = factor(
  #     n_events_cum > 0, 
  #     levels = c(FALSE, TRUE), 
  #     labels = c("0 Events", "1+ Events")
  #     )
  # )
```






# Table 1: Demographics

<!-- School-level student count, attendance, and percentages appear  -->
<!-- non-normally distributed. Plan to summarise by non-parametric statistics. -->

```{r}
# t1a_df |> 
#   ggplot(aes(value, color = data_schoolyear)) +
#   geom_density() +
#   theme(legend.position = "none") +
#   facet_wrap(. ~ variable, scales = "free")
```

<!-- Within-school changes over time are more symmetrically distributed.  -->
<!-- Several change values appear centered above zero, suggesting mean  -->
<!-- increases in reported numbers over time.  -->

```{r}
# t1a_df |> 
#   filter(data_schoolyear != baseline_schoolyear) |> 
#   ggplot(aes(abs_change, color = data_schoolyear)) +
#   geom_density() +
#   theme(legend.position = "none") +
#   geom_vline(xintercept = 0, linetype = "dashed") +
#   facet_wrap(. ~ variable, scales = "free")
```

Calculate, summarize, and hypothesis-test changes in each variable from 
baseline (2021-2022) to each subsequent year.


```{r}
# # count number of events (2023-2024 SY only)
# school_time_invariant <- full_join(
#   school_df |>
#         distinct(school_id, grade_level, region, region_3grp, wave),
#   use_counts |>
#     pivot_wider(
#     )
#     filter(data_schoolyear == "2023-2024") |> 
#     summarise(n_events = n(), .by = c(school_id)),
#     by = c("school_id")
# ) |> 
#   mutate(
#     n_schools   = 1,
#     n_events    = coalesce(n_events, 0),
#     any_events_2l = factor(n_events > 0, 
#          levels = c(FALSE, TRUE),
#          labels = c("0 Events",	"1+ Events")
#          ),
#     any_events_3l = factor(
#       case_when(
#         wave == "Fall 2024" ~ "NA - Phase 4",
#         n_events == 0 ~ "0 Events",
#         n_events > 0 ~ "1+ Events"
#       )
#     )
#   )
```


```{r}

```


Build summary tables:

```{r}
# school-level (time-invariant) chars
t1a <- tableby(
  data_schoolyear ~ wave + grade_level + n_students + region + region_3grp,
  data = read_analytic(params$analytic_files$school_counts_annual),
  test = FALSE,
  total = FALSE,
  control = tableby.control(
    numeric.stats = c("medianq1q3", "range")
  )
  ) |> 
  summary(text = TRUE) |> 
  as.data.frame() 


# School-level student counts (annual):
(t1b <- tableby(
    data_schoolyear ~ 
      p_student_race_american_indian + 
      p_student_race_asian_hawaiian_pacific_islander + 
      p_student_race_black_non_hispanic + 
      p_student_race_hispanic + 
      p_student_race_middle_eastern_or_north_african +
      p_student_race_multi + 
      p_student_race_white_non_hispanic + 
      p_student_race_missing +
      
      p_student_gender_female + 
      p_student_gender_male + 
      p_student_gender_non_binary +
      p_student_gender_missing + 
      
      p_esl_yes + p_frm_yes + p_homelessness_yes +
      
      attendance + p_asthma_yes + p_food_allergy_yes + p_non_food_allergy_yes +
      
      p_student_504_yes + p_iep_yes, # + any_events_3l,
  data = read_analytic(params$analytic_files$school_counts_annual),
  test = FALSE,
  total = FALSE,
  control = tableby.control(
    numeric.stats = c("medianq1q3", "range")
  )
  ) |> 
  summary(text = TRUE) |> 
  as.data.frame() )
```



# Bivariate Tests

  * for each variable, calculate year-vs-baseline hypothesis tests (paired t-tests)
    
<!-- * data review, processing, recoding -->
<!-- * Check cell counts for all groups (small groups, e.g. NB) -->


    

```{r}
# t-tests for change from baseline to each subsequent year in student counts
# (H0: mean within-school change = 0; Ha: change != 0) 
read_analytic(params$analytic_files$school_counts_annual) |> 
  select(school_id, data_schoolyear, attendance, starts_with("p_")) |> 
  select(-any_of(c(
    "p_students",
    "p_student_race_middle_eastern_or_north_african"
    ))) |> 
  pivot_longer(c("attendance", starts_with("p")), names_to = "var") |> 
  pivot_wider(
    names_from = "data_schoolyear"
  ) |> 
  
  mutate(
    across(
      all_of(c("2022-2023", "2023-2024", "2024-2025")),
      ~ .x - `2021-2022`
      )
    ) |> 
  
  pivot_longer(
    c("2022-2023", "2023-2024", "2024-2025"),
    names_to = "data_schoolyear",
    values_to = "change_from_bl"
    ) |> 
  filter(!is.na(change_from_bl)) |> 
  
  summarise(
    h0        = "H0: No mean within-school change from 2021-2022 baseline",
    h1        = "H1: Mean within-school change from 2021-2022 baseline != 0",
    n_schools = n(),
    tt        = broom::tidy(t.test(change_from_bl)),
    .by       = c(var, data_schoolyear)
  ) |> 
  unnest(tt) |> 
  mutate(p.value.adj = p.adjust(p.value, method = "holm"),
         .by = var)
```

```{r}
# combine tables and write to file
list(t1a, t1b) |> 
  map_dfr(rename, x = 1) |> 
    
  openxlsx::write.xlsx(
    here::here("a11_sow3_longitudinal", "output", Sys.Date(), "t1_characteristics.xlsx"),
    asTable = TRUE
  )

```



## Table 2: Utilization, 2023-2024 School Year


```{r}
(t2 <- c(
  list(levels(read_analytic(params$analytic_files$school_counts_annual)$any_events_3l)), 
  combn(levels(read_analytic(params$analytic_files$school_counts_annual)$any_events_3l), 2, simplify = FALSE)
  ) |> 
  (\(.) set_names(., map(., paste, collapse = " vs. ")))() |> 
  map(~{
    tableby(
  any_events_3l ~ notest(wave) + grade_level + n_students + region + region_3grp + p_student_race_american_indian + 
      p_student_race_asian_hawaiian_pacific_islander + 
      p_student_race_black_non_hispanic + 
      p_student_race_hispanic + 
      p_student_race_middle_eastern_or_north_african +
      p_student_race_multi + 
      p_student_race_white_non_hispanic + 
      p_student_race_missing +
      
      p_student_gender_female + 
      p_student_gender_male + 
      p_student_gender_non_binary +
      p_student_gender_missing + 
      
      p_esl_yes + p_frm_yes + p_homelessness_yes +
      
      attendance + p_asthma_yes + p_food_allergy_yes + p_non_food_allergy_yes +
      
      p_student_504_yes + p_iep_yes,
  data = read_analytic(params$analytic_files$school_counts_annual) |> 
    filter(data_schoolyear == "2023-2024") |> 
    filter(any_events_3l %in% .x),
  control = tableby.control(
    numeric.stats = c("medianq1q3", "range")
  )
  
) |> 
  
  summary(text = TRUE) |> 
  as.data.frame()
    
  }))
```


```{r}
(t2[[1]] |> 
  rename(x = 1) |> 
  select(-matches("p.value")) |> 
  
  full_join(
    t2 |> 
    map_dfr(
      ~ .x |>
        janitor::clean_names() |>
        select(x, p_value) |>
        filter(str_length(p_value) > 0),
      .id =  "comparison"
      ) |> 
    
    pivot_wider(
      names_from = comparison,
      names_prefix = "pval ",
      values_from = p_value
    ) ,
    
    by = "x"
    
  )) |> 
  
  openxlsx::write.xlsx(
    here::here("a11_sow3_longitudinal", "output", Sys.Date(), "t2_utilization.xlsx"),
    asTable = TRUE
  )

```




```{r}
# Table 2b: Utilization Detail
t2_df <- right_join(
    read_analytic(params$analytic_files$school_chars),
    read_analytic(params$analytic_files$inhaler_events),
    by = "school_id"
    )
```


```{r}
tableby(
  data_schoolyear ~
    # recipient
    rcp_sex + rcp_race + rcp_eth + rcp_age +
    rcp_role + sym_new + rcp_dx_pre + inh_use_prev +

    # incident
    season + why_given +

    # school chars
    wave + grade_level +

    # disposition
    hypo_disp_rcd + true_disp_rcd,
  
  test = FALSE,
  data = t2_df |> filter(data_schoolyear %in% c("2023-2024", "2024-2025")) |> droplevels(),
  total = FALSE
  ) |>
  summary(text = TRUE, labelTranslations = list(rcp_age = "Age")) |>
  as.data.frame()
  # write_output("Table2_Utilization")
```


## Attendance, Asthma, IEPs, 504s


# Longitudinal Models

* Check for significant interactions:
    * time by gender (504 outcome)
    * time by gender (IEP outcome)
    * time by gender (asthma outcome)
    * time by race (504 outcome)
    * time by race (IEP outcome)
    * time by race (asthma outcome)

```{r}
read_analytic(params$analytic_files$school_counts_annual)
```


```{r}
t3_df <- read_analytic(params$analytic_files$student_counts)

gender_probs <- t3_df |> 
  summarise(
    n   = sum(student_count),
    .by = student_gender
    ) |> 
  na.omit() |> 
  mutate(p = n / sum(n), .keep = "unused")

t3_df_imp <- merge(
  gender_probs,
  t3_df |> 
    filter(is.na(student_gender)) |> 
    select(-student_gender),
  by = NULL
  ) |> 
  mutate(across(
    c(matches("_no$"), matches("_yes$"), 
      matches("_present$"), matches("_absent$")),
    ~ .x*p
    )) |> 
  
  select(-p) |> 
  rbind(
    t3_df |> 
      filter(!is.na(student_gender))
    ) |> 
  
    mutate(
      student_gender = fct_na_value_to_level(
        student_gender, "Unknown"
        ),
      student_race = fct_collapse(fct_na_value_to_level(
        student_race, "Unknown"
        ),
        "Multi, Other, Unknown" = c("Multi", "Unknown", "Middle Eastern or North African")
      )
    ) |> 
    droplevels()
```


```{r}
# nrow(t3_df)
# nrow(t3_df_imp)
# 
# cbind(
#   colSums(t3_df |> select(where(is.numeric))),
#   colSums(t3_df_imp |> select(where(is.numeric)))
# )
```


```{r}
# create dataset with asthma counts by race, gender, school, schoolyear
t3_df_list <- list(
  asthma = t3_df_imp |> 
    arrange(school_id, data_schoolyear) |> 
    summarise(
      y_var   = "asthma",
      n_y_yes = round(sum(asthma_count_yes, na.rm = TRUE)),
      n_y_no  = round(sum(asthma_count_no, na.rm = TRUE)),
      .by     = c(school_id, data_schoolyear, data_date, student_gender, student_race)
    )
)

# create dataset with iep counts by asthma, race, gender, school, schoolyear
t3_df_list[["iep"]] <- t3_df_imp |> 
    arrange(school_id, data_schoolyear) |> 
    summarise(
      y_var = "iep",
      n_y_yes = round(sum(iep_count_yes, na.rm = TRUE)),
      n_y_no  = round(sum(iep_count_no, na.rm = TRUE)),
      .by   = c(school_id, data_schoolyear,  data_date, asthma_status_lab, student_gender, student_race)
    )

# create dataset with 504 counts by asthma, race, gender, school, schoolyear
t3_df_list[["x504"]] <- t3_df_imp |> 
    arrange(school_id, data_schoolyear) |> 
    summarise(
      y_var = "x504",
      n_y_yes = round(sum(student_504_count_yes, na.rm = TRUE)),
      n_y_no  = round(sum(student_504_count_no, na.rm = TRUE)),
      .by   = c(school_id, data_schoolyear, data_date, asthma_status_lab, student_gender, student_race)
    ) 

# create dataset with attendance counts by asthma, race, gender, school, schoolyear
t3_df_list[["attendance"]] <- t3_df_imp |> 
    arrange(school_id, data_schoolyear) |> 
    summarise(
      y_var = "days_present",
      n_y_yes = round(sum(student_days_present, na.rm = TRUE)),
      n_y_no  = round(sum(student_days_absent, na.rm = TRUE)),
      .by   = c(school_id, data_schoolyear, data_date, asthma_status_lab, student_gender, student_race)
    )
```



```{r}
# combine with school and utilization data
t3_df_list_mod <- map(t3_df_list, ~{
  full_join(
    read_analytic(params$analytic_files$school_chars), .x, by = c("school_id")) |> 
  left_join(
    read_analytic(params$analytic_files$inhaler_event_counts), 
    by = c("school_id",  "data_schoolyear")) |>
    arrange(school_id, data_date) |> 
    mutate(
      
      data_schoolyear = fct_inorder(data_schoolyear),
      data_schoolyear_seq = as.numeric(data_schoolyear),
      data_schoolyear_num = data_schoolyear_seq + 2021,
     active_months = round(as.numeric(
       difftime(data_date, impl_date, units = "days")
       )/(30.44)),
     active_months = case_when(
       active_months > 0 ~ active_months,
       .default = 0
       ),
     .before = 1
     ) |> 
    filter(!is.na(y_var)) |> 
    mutate(across(where(is.character), factor))  |> 
    
    filter(!(n_y_yes + n_y_no) == 0)
  })
```



```{r}
# check counts by race:
t3_df_list_mod |> 
  map(~{
    .x |> 
      reframe(
        level = c("yes", "no"),
        value = c(sum(n_y_yes),sum(n_y_no)),
        .by = c("data_schoolyear", "student_race")
      ) |> 
      pivot_wider(
        names_from = data_schoolyear
      ) |> 
      arrange(level, desc(`2021-2022`))
  })
```


```{r}
# check counts by gender
t3_df_list_mod |> 
  map(~{
    .x |> 
      reframe(
        level = c("yes", "no"),
        value = c(sum(n_y_yes),sum(n_y_no)),
        .by = c("data_schoolyear", "student_gender")
      ) |> 
      pivot_wider(
        names_from = data_schoolyear
      ) |> 
      arrange(level, desc(`2021-2022`))
  })
```


```{r}
write_rds(
  t3_df_list_mod,
  find_dated_output("gee_mod_data_list.rds")
)
```


Model each outcome probability by school year, race, gender, 
active_months, wave, grade_level, utilization

Counts



```{r}
# map2(t3_mod_list, t3_df_list_mod, function(m, d) {
#     lmtest::lrtest(
#       update(m, data = d),
#       update(m, . ~ . + data_schoolyear:student_race, data = d)
#     )
#   })
# lmtest::lrtest(t3_mod_list[[1]])

```


<!-- Asthma:  -->

```{r}
f_list <- c(
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(asthma_status_lab + student_race + student_gender) +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_race + student_gender) + asthma_status_lab +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(asthma_status_lab + student_race) + student_gender +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(asthma_status_lab + student_gender) + student_race +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(asthma_status_lab) + student_gender + student_race +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_gender) + asthma_status_lab + student_race +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_race) + asthma_status_lab + student_gender +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear + student_race + asthma_status_lab + student_gender +
      active_months + any_events_cum +
      wave + grade_level + region_3grp
)


f_list_2 <- c(
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_race + student_gender) +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_gender) + student_race +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear*(student_race) + student_gender +
      active_months + any_events_cum +
      wave + grade_level + region_3grp,
   cbind(n_y_yes, n_y_no) ~
     data_schoolyear + student_race + student_gender +
      active_months + any_events_cum +
      wave + grade_level + region_3grp
)
```


```{r}
if(!file.exists(find_dated_output("gee_mod_search.rds"))) {
  c(
    rlist::list.expand(
      d = c("iep", "x504", "attendance"), 
      f = f_list
      ),
    rlist::list.expand(
      d = c("asthma"), 
      f = f_list_2
      )
    ) |> 
    map(~{
      eval(bquote(
        geepack::geeglm(
          .(.x$f),
          family = binomial(link = "logit"),
          data   = droplevels(na.omit(t3_df_list_mod[[.(.x$d)]])),
          id     = factor(school_id),
          waves  = data_schoolyear_seq,
          corstr = "independence"
          )
      ))
      }) |> 
  
  write_rds(find_dated_output("gee_mod_search.rds"))
}

t3_mod_search <- read_rds(find_dated_output("gee_mod_search.rds"))
```

```{r}
if(!file.exists(find_dated_output("gee_mod_search_fitstats.rds"))) {
  t3_mod_search_fitstats <- map_dfr(t3_mod_search, ~{
    data.frame(
      call = paste(as.character(.x$call), collapse = "; "),
      data.frame(as.list(geepack::QIC(.x)))
    )
  })
  write_rds(
    t3_mod_search_fitstats,
    find_dated_output("gee_mod_search_fitstats.rds")
  )
}

t3_mod_search_fitstats <- read_rds(find_dated_output("gee_mod_search_fitstats.rds"))
```


```{r}
# cbind(
#   c(
#     rlist::list.expand(
#       d = c("iep", "x504", "attendance"), 
#       f = f_list
#       ),
#     rlist::list.expand(
#       d = c("asthma"), 
#       f = f_list_2
#       )
#     ) |> 
#   map_dfr(~{data.frame(d = .x$d, f = as.character(.x$f)[3])}),
#   
#   t3_mod_search_fitstats
# ) |>
#   mutate(
#     intx = gsub(".*\\(([^\\)]+)\\).*", "\\1", f),
#     intx = case_when(grepl("schoolyear", intx) ~ "(none)", .default = intx)
#     ) |> 
#   group_split(d) |> 
#   map(~{
#     .x |> 
#       ggplot(aes(intx, QICu)) +
#     geom_point() +
#       coord_flip()
#     
#   })
```

```{r}
cbind(
  c(
    rlist::list.expand(
      d = c("iep", "x504", "attendance"), 
      f = f_list
      ),
    rlist::list.expand(
      d = c("asthma"), 
      f = f_list_2
      )
    ) |> 
  map_dfr(~{data.frame(d = .x$d, f = as.character(.x$f)[3])}),
  t3_mod_search_fitstats
  ) |> 
  filter((QICu - min(QICu) <= 6), .by = d) |> 
  arrange(d, QICu)
```




```{r}
# if(!file.exists(find_dated_output("gee_mod_select.rds"))) {
#   
#   c(
#     rlist::list.expand(
#         d = c("iep", "x504", "attendance"), 
#         f = cbind(n_y_yes, n_y_no) ~ 
#           data_schoolyear * (asthma_status_lab) + student_gender + student_race + 
#           active_months + any_events_cum + wave + grade_level + region_3grp
#         ),
#       rlist::list.expand(
#         d = c("asthma"), 
#         f = cbind(n_y_yes, n_y_no) ~ 
#           data_schoolyear + student_gender + student_race + 
#           active_months + any_events_cum + wave + grade_level + region_3grp
#         )
#       ) |> 
#       map(~{
#         eval(bquote(
#           geepack::geeglm(
#             .(.x$f),
#             family = binomial(link = "logit"),
#             data   = droplevels(na.omit(t3_df_list_mod[[.(.x$d)]])),
#             id     = factor(school_id),
#             waves  = data_schoolyear_seq,
#             corstr = "independence"
#             )
#         ))
#         }) |> 
#     
#     write_rds(find_dated_output("gee_mod_select.rds"))
# }
# 
# t3_mod_list <- read_rds(find_dated_output("gee_mod_select.rds"))
```


```{r}
# rlist::list.expand(
#   d = c("iep", "x504", "attendance"), 
#   f = f_list
#   ) |> 
#   map(~{
#     eval(bquote(
#       geepack::geeglm(
#         .(.x$f),
#         family = binomial(link = "logit"),
#         data   = droplevels(na.omit(t3_df_list_mod[[.(.x$d)]])),
#         id     = factor(school_id),
#         waves  = data_schoolyear_seq,
#         corstr = "independence"
#         )
#     ))
#     })
```




```{r}
# broom::tidy(t3_mod_list[[1]])
# 
# emmeans::emmeans(
#   t3_mod_list[[1]],
#   ~ data_schoolyear | student_gender,
#   type = "response"
#   ) |>
#   emmeans::contrast("trt.vs.ctrl")
```


```{r}
broom::tidy(geepack::geeglm(
  cbind(n_y_yes, n_y_no) ~
    data_schoolyear*(asthma_status_lab + student_gender + student_race) + 
          active_months + any_events_cum + wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = droplevels(na.omit(t3_df_list_mod[["x504"]])),
  id     = factor(school_id),
  waves  = data_schoolyear_seq,
  corstr = "independence"
  ))
```

```{r}
geemod_x504 <- geepack::geeglm(
  cbind(n_y_yes, n_y_no) ~
    data_schoolyear*(asthma_status_lab + student_gender + student_race) + 
          active_months + any_events_cum + wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = droplevels(na.omit(t3_df_list_mod[["x504"]])),
  id     = factor(school_id),
  waves  = data_schoolyear_seq,
  corstr = "independence"
  )
```


504, compare asthma status by year

```{r}
emmeans::emmeans(
  geemod_x504, revpairwise ~ asthma_status_lab | data_schoolyear,
  type = "response"
  )
```

504, compare year vs. baseline by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, trt.vs.ctrl ~ data_schoolyear | asthma_status_lab,
  type = "response"
  )
```

504, compare year vs. previous by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, consec ~ data_schoolyear | asthma_status_lab,
  type = "response"
  )
```

504, compare race/ethnicity status by year

```{r}
emmeans::emmeans(
  geemod_x504, revpairwise ~ student_race | data_schoolyear,
  type = "response"
  )
```

504, compare year vs baseline by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, trt.vs.ctrl ~ data_schoolyear | student_race,
  type = "response"
  )
```

504, compare year vs previous by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, consec ~ data_schoolyear | student_race,
  type = "response"
  )
```



504, compare asthma status by year

```{r}
emmeans::emmeans(
  geemod_x504, revpairwise ~ asthma_status_lab | data_schoolyear,
  type = "response"
  )
```

504, compare year vs. baseline by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, trt.vs.ctrl ~ data_schoolyear | asthma_status_lab,
  type = "response"
  )
```

504, compare year vs. previous by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, consec ~ data_schoolyear | asthma_status_lab,
  type = "response"
  )
```

504, compare race/ethnicity status by year

```{r}
emmeans::emmeans(
  geemod_x504, revpairwise ~ student_gender | data_schoolyear,
  type = "response"
  )
```

504, compare year vs baseline by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, trt.vs.ctrl ~ data_schoolyear | student_gender,
  type = "response"
  )
```

504, compare year vs previous by asthma status

```{r}
emmeans::emmeans(
  geemod_x504, consec ~ data_schoolyear | student_gender,
  type = "response"
  )
```






```{r}
summary(geepack::geeglm(
  cbind(n_y_yes, n_y_no) ~
    asthma_status*(factor(data_schoolyear) + student_race + student_gender) +

    active_months + n_events_cum + 
    
    wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = na.omit(t3_df_list_mod[["asthma"]]),
  id     = factor(school_id),
  waves  = data_schoolyear_seq,
  corstr = "independence"
  ))
```




```{r}
add_todo(paste(
  "Consider addl interactions. Possible to collapse race and gender?",
  "Currently rank defficient if including all student-level 2-way",
  "interactions with year."
  ))

summary(t3a_mod <- geepack::geeglm(
  cbind(student_days_present, student_days_absent) ~
    asthma_status*(factor(data_schoolyear) + student_race + student_gender) +

    active_months + wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = na.omit(t3a_df),
  id     = factor(school_id),
  waves  = data_schoolyear,
  corstr = "independence"
  ))
```


```{r}
t3a_est <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::emmeans(
      t3a_mod,
      ~ data_schoolyear | asthma_status,
      by = .x,
      type = "response",
      weights = "proportional"
    )
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)

t3a_or <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::contrast(emmeans::emmeans(
      t3a_mod,
      ~ data_schoolyear | asthma_status,
      by = .x,
      weights = "proportional"
    ), "trt.vs.ctrl", type = "response", infer = TRUE)
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)


# -------------------------------------------------------------------------

t3b_df <- cps_df |>
  select(-student_count) |>
  pivot_longer(
    c("iep_count_yes", "iep_count_no"),
    names_to = "iep_status",
    values_to = "student_count"
    ) |>
  mutate(
    student_days_expected = student_count*cps_base_days,
    student_days_present = student_days_expected*(average_attendance/100),
    student_days_absent  = student_days_expected - student_days_present,
    .after               = average_attendance
  ) |>
  summarise(
    n_students         = sum(student_count),
    average_attendance = weighted.mean(average_attendance, student_count),
    across(starts_with("student_days"), ~ round(sum(.x))),
    .by = c(
      # school level
      school_id, wave, grade_level, region, region_2grp, region_3grp,
      pilot, impl_year, impl_date, data_schoolyear, data_date, active_months,

      # student level
      iep_status, student_gender, student_race
    )
  ) |>

  droplevels()

summary(t3b_mod <- geepack::geeglm(
  cbind(student_days_present, student_days_absent) ~
    iep_status*(factor(data_schoolyear) + student_race + student_gender) +

    active_months + wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = na.omit(t3b_df),
  id     = factor(school_id),
  waves  = data_schoolyear,
  corstr = "independence"
  ) |>
  write_data("t3b_mod_attd_iep"))



t3b_est <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::emmeans(
      t3b_mod,
      ~ data_schoolyear | iep_status,
      by = .x,
      type = "response",
      weights = "proportional"
    )
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)

t3b_or <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::contrast(emmeans::emmeans(
      t3b_mod,
      ~ data_schoolyear | iep_status,
      by = .x,
      weights = "proportional"
    ), "trt.vs.ctrl", type = "response", infer = TRUE)
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)





# -------------------------------------------------------------------------


t3c_df <- cps_df |>
  select(-student_count) |>
  pivot_longer(
    c("student_504_count_yes", "student_504_count_no"),
    names_to = "x504_status",
    values_to = "student_count"
  ) |>
  mutate(
    student_days_expected = student_count*cps_base_days,
    student_days_present = student_days_expected*(average_attendance/100),
    student_days_absent  = student_days_expected - student_days_present,
    .after               = average_attendance
  ) |>
  summarise(
    n_students         = sum(student_count),
    average_attendance = weighted.mean(average_attendance, student_count),
    across(starts_with("student_days"), ~ round(sum(.x))),
    .by = c(
      # school level
      school_id, wave, grade_level, region, region_2grp, region_3grp,
      pilot, impl_year, impl_date, data_schoolyear, data_date, active_months,

      # student level
      x504_status, student_gender, student_race
    )
  ) |>

  droplevels()

summary(t3c_mod <- geepack::geeglm(
  cbind(student_days_present, student_days_absent) ~
    x504_status*(factor(data_schoolyear) + student_gender) +
  student_race  +
    active_months + wave + grade_level + region_3grp,
  family = binomial(link = "logit"),
  data   = na.omit(t3c_df),
  id     = factor(school_id),
  waves  = data_schoolyear,
  corstr = "independence"
  ) |>
  write_data("t3c_mod_attd_504"))


t3c_est <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::emmeans(
      t3c_mod,
      ~ data_schoolyear | x504_status,
      by = .x,
      type = "response",
      weights = "proportional"
    )
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)

t3c_or <- set_names(c("student_gender", "student_race")) |>
  map(~{
    emmeans::contrast(emmeans::emmeans(
      t3c_mod,
      ~ data_schoolyear | x504_status,
      by = .x,
      weights = "proportional"
    ), "trt.vs.ctrl", type = "response", infer = TRUE)
  }) |>
  map_dfr(data.frame, .id = "byvar") |>
  mutate(level = coalesce(student_gender, student_race),
         .keep = "unused",
         .after = byvar)

# -------------------------------------------------------------------------





t3_ests <- list(
  t3a_est,
  t3b_est,
  t3c_est
  ) |>

  map_dfr(~{
    pivot_longer(
      .x,
      matches("status"), values_to = "status", names_to = "strata")
  }) |>
  pivot_wider(
    names_from = data_schoolyear,
    values_from = c(prob, lower.CL, upper.CL),
    id_cols = c(strata, status, byvar, level)
  ) |>
  relocate(
    matches("2022"), matches("2023"), matches("2024"), matches("2025"),
    .after = level
    )



t3_ors <- list(
  t3a_or,
  t3b_or,
  t3c_or
) |>
  map_dfr(~{
    pivot_longer(
      .x,
      matches("status"), values_to = "status", names_to = "strata")
  }) |>
  pivot_wider(
    names_from = contrast,
    values_from = c(odds.ratio, lower.CL, upper.CL, p.value),
    id_cols = c(strata, status, byvar, level)
  ) |>
  relocate(
    matches("2023"), matches("2024"), matches("2025"),
    .after = level
  )

t3 <- full_join(
  t3_ests,
  t3_ors
  ) |>
  arrange(fct_inorder(strata), desc(status)) |>

  write_output("Table3_Atttendance")

# -------------------------------------------------------------------------




# Table 4: Asthma
t4a_df <- cps_df |>
  summarise(
    student_count = sum(student_count),
    .by = c(
      school_id,
      data_schoolyear, impl_year, impl_date, active_months, #util,
      student_race, student_gender, asthma_status_lab
    )) |>
  pivot_wider(
    names_from = asthma_status_lab,
    values_from = student_count,
    id_cols = c(
      school_id,
      data_schoolyear, impl_year, impl_date, active_months, #util,
      student_race, student_gender
      ),
    values_fill = 0
    ) |>
  mutate(
    util   = fct_relevel(case_when(
      school_id %in% inh_df$school_id ~ "1+",
      .default = "None"
    ), "None"),
    student_race = fct_relevel(student_race, "Black Non-Hispanic"),
    wave         = fct_relevel(as.character(impl_date), "2024-09-01")
    ) |>
  droplevels()


summary(t4_mod <- geepack::geeglm(
  cbind(`Asthma Dx`, `No Asthma Dx`) ~
    student_race +
    student_gender +
    factor(data_schoolyear)*(

    wave +
    util),
  family = binomial(link = "logit"),
  data   = na.omit(t4a_df),
  id     = factor(school_id),
  waves  = data_schoolyear,
  corstr = "independence"
))


t4a_est <- set_names(c("student_gender", "student_race", "util", "wave")) |>
    map(~{
      emmeans::emmeans(
        t4_mod,
        ~ data_schoolyear,
        by = .x,
        type = "response",
        weights = "proportional"
      )
    }) |>
    map_dfr(data.frame, .id = "byvar") |>
    mutate(level = coalesce(student_gender, student_race, util, wave),
           .keep = "unused",
           .after = byvar)


t4b_or <- set_names(c("student_gender", "student_race", "util", "wave")) |>
    map(~{
      emmeans::contrast(emmeans::emmeans(
        t4_mod,
        ~ data_schoolyear,
        by = .x,
        weights = "proportional"
      ),
      "trt.vs.ctrl",
      type = "response",
      infer = TRUE
      )
    }) |>
    map_dfr(data.frame, .id = "byvar") |>
    mutate(level = coalesce(student_gender, student_race, util, wave),
           .keep = "unused",
           .after = byvar)


t4 <- full_join(


  t4a_est |>
    pivot_wider(
      names_from = data_schoolyear,
      values_from = c(prob, lower.CL, upper.CL),
      id_cols = c(byvar, level)
    ) |>
    relocate(
      matches("2022"), matches("2023"), matches("2024"), matches("2025"),
      .after = level
    ),


  t4b_or |>
    pivot_wider(
      names_from = contrast,
      values_from = c(odds.ratio, lower.CL, upper.CL, p.value),
      id_cols = c(byvar, level)
    ) |>
    relocate(
      matches("2023"), matches("2024"), matches("2025"),
      .after = level
    )

) |>
  write_output(("Table4_Asthma"))

# Utilization data --------------------------------------------------------







# # Read and modify data
# if(!exists("inh_use_raw")) {
#     inh_use_raw <- boxr::box_read_excel(
#         file_id = get_global("raw_files")$utilization$box_id,
#         sheet   = get_global("raw_files")$utilization$sheet
#     )
# }
#
# inh_use <- inh_use_raw |>
#     filter(!is.na(ID)) |>
#     drop_empty_cols() |>
#     detect_factors(titlize = FALSE) |>
#     rename_with(str_trim) |>
#     set_colnames_as_labels() |>
#     (\(.)
#      . |> set_names(sprintf("x%03d", 1:ncol(.)))
#      )()





# # match inhaler use to school names and ids
# inh_use <- inh_use |>
#
#     rename_by_label(
#         label = "At what school was the stock inhaler given?",
#         newname = "school_name"
#         ) |>
#     mutate(
#         school_match_name = gsub("_", "", snakecase::to_snake_case(school_name)),
#         school_match_name = case_when(
#             school_match_name  == "kenwood" ~ "kenwoodhs",
#             school_match_name  == "lindbloom" ~ "lindblomhs",
#             school_match_name  == "cant" ~ "canty",
#             .default = school_match_name
#         ),
#         .after = 1
#         ) |>
#
#     mutate(across(where(is.POSIXt), as.Date)) |>
#
#     left_join(
#         school_names_ids |>
#             distinct(school_match_name, school_id),
#         by = "school_match_name"
#     ) |>
#
#     write_data(get_global("analytic_files")$inh_use, csv = TRUE)




# Apply factor codes from dataframe/list
# factor_codes = list(
#   vname_1 = data.frame(
#     levels = c(1, 2, 3),
#     labels = c("a", "b", "c")
#     )
#   )
# my_cleaned_data <- my_raw_data
# for (i in intersect(names(factor_codes), names(my_raw_data))) {
#   my_cleaned_data[[i]] <- factor(
#     my_raw_data[[i]],
#     levels = factor_codes[[i]]$level,
#     labels = factor_codes[[i]]$label
#   )
# }

##########################################################################
# Save output
# out_tables[["My Data"]] <- my_data |>
#     write_data("my_data")







inh_use |>
    rename_by_label(label = "ID", newname = "row_id") |>
    rename_by_label("Recipient Sex",  "rcp_sex") |>
    rename_by_label("Recipient Race", "rcp_race") |>
    rename_by_label("Recipient Ethnicity", "rcp_eth") |>
    rename_by_label("Recipient age (If recipient is a visitor over the age of 18, student teacher, or school personnel select \"Adult\")", "rcp_age") |>
    rename_by_label("On what date the stock inhaler given?", "admin_date") |>
    rename_by_label("Location of Incident (Playground, gym, nurse's office, specific classroom number, etc.)", "adm_loc") |>
    rename_by_label("Under what protocol was the stock inhaler given?", "adm_protocol") |>
    rename_by_label("Was the school nurse present or notified?", "nurse_present") |>
    rename_by_label("Exposure to pollen allergen", "exp_pollen") |>
    rename_by_label("Exposure to dust allergen", "exp_dust") |>
    rename_by_label("Exposure to animal dander", "exp_anim") |>
    rename_by_label("Exposure to smoke in the environment", "exp_smok") |>
    rename_by_label("Exposure to air pollution in the environment", "exp_air") |>
    rename_by_label("Exposure to hot weather", "exp_heat") |>
    rename_by_label("Exposure to cold weather", "exp_cold") |>
    rename_by_label("Physical activity or exercise", "exp_phys") |>
    rename_by_label("Unknown", "exp_unk") |>
    rename_by_label("No exposure", "exp_none") |>
    rename_by_label("Other...29", "exp_other") |>
    rename_by_label("Shortness of breath", "sym_sob") |>
    rename_by_label("Wheezing", "sym_wheeze") |>
    rename_by_label("Coughing", "sym_cough") |>
    rename_by_label("Chest tightness", "sym_tight") |>
    rename_by_label("Rapid breathing", "sym_rapid") |>
    rename_by_label("Difficulty speaking or completing sentences", "sym_speak") |>
    rename_by_label("Chest pain", "sym_chpain") |>
    rename_by_label("Flaring of the nostrils", "sym_flare") |>
    rename_by_label("Retraction of neck and chest muscles", "sym_musc") |>
    rename_by_label("Bluish or pale coloration of the lips or fingertips", "sym_blue") |>
    rename_by_label("Anxiety or restlessness", "sym_anx") |>
    rename_by_label("breathing difficulty", "sym_brdiff") |>
    rename_by_label("No symptoms", "sym_none") |>
    rename_by_label("Other...46", "sym_other") |>
    rename_by_label("Correct Pathway", "pathway_true") |>
    rename_by_label("Chosen right pathway?", "pathway_chosen") |>
    rename_by_label("What is the name of the inhaler/medication given? (HFA Albuterol, ProAir RespiClick, etc.)", "inh_med") |>
    rename_by_label("Was a spacer device used? (Either plastic or paper)", "spacer") |>
    rename_by_label("How many puffs were administered using the HFA Inhaler?", "n_puffs_hfa") |>
    rename_by_label("Compliant with nurse-selected pathway", "pathway_chosen_compl") |>
    rename_by_label("compliant with correct pathway", "pathway_true_compl") |>
    rename_by_label("Was a disposable mouthpiece used?", "mouthpiece") |>
    rename_by_label("How many puffs were administered using the ProAir RespiClick?", "n_puffs_respi") |>
    rename_by_label("Type of person administered", "rcp_role") |>
    rename_by_label("Previous Dagnosis", "rcp_dx_pre") |>
    rename_by_label("Has the student since received an asthma diagnosis?", "rcp_dx_post") |>
    rename_by_label("Do they need assistance connecting with a healthcare provider or finding insurance coverage?", "hc_help") |>
    rename_by_label("Any Previous Usage", "inh_use_prev") |>
    rename_by_label("Was inhaler administered to prevent symptoms or given due to breathing difficulties?", "why_given") |>
    rename_by_label("According to the protocol, was the individual exhibiting mild-moderate symptoms or severe symptoms?", "severity") |>
    rename_by_label("As far as you are aware, is this the first time the individual has ever experienced these kind of symptoms at school?", "sym_new") |>
    rename_by_label("In your professional opinion, what would have happened if the student/individual did not have access to stock inhaler medication?", "hypo_disp") |>
    rename_by_label("Hypothetical_Disposition_simplified", "hypo_disp_rcd") |>
    rename_by_label("What DID happen to the student/individual after the inhaler was given?", "true_disp") |>
    rename_by_label("Actual_Disposition_simplified", "true_disp_rcd") |>
    rename_by_label("Did the individual go to the hospital after emergency services came?", "hosp") |>
    rename_by_label("How did they get to the hospital?", "transport") |>
    relocate(-starts_with("x"))
school_names_ids <- list.files(file.path(params$project_dir, "data-raw"), "xls", full.names = TRUE) |>
    set_names() |>
    map(~{
        nms <- readxl::read_excel(.x, sheet = "Schools", col_names = FALSE, n_max = 4) |>
            filter(!grepl("\\*", ...1)) |>
            slice(1:2) |>
            summarise(across(everything(), ~ paste(na.omit(unlist(.x)), collapse = "; "))) |>
            unlist()

        readxl::read_excel(.x, sheet = "Schools", col_names = nms)
    }) |>
    map_dfr(~{
        .x |>
            janitor::clean_names() |>
            select(matches("school_id"), school_name, total) |>
            set_names(c("school_id", "school_name", "total"))
    }, .id = "source") |>
    mutate(
        source = basename(source),
        school_match_name = gsub("_", "", snakecase::to_snake_case(school_name))
        ) |>
    filter(!is.na(school_id)) |>
    filter(!grepl("school", tolower(school_id)))











# TODO: investigate, deal with missing data

```





```{r}
gee_a <- read_rds(
  file.path(
    params$project_dir,
    "data", Sys.Date(), 
    "gee_mod_list.rds"
))
```

```{r}
# gee_b <- read_rds(
#   file.path(
#     params$project_dir,
#     "data", Sys.Date(), 
#     "gee_mod_asth_intx_list.rds"
# ))
```



# Asthma Outcome

Model summary

```{r}
mod_current <- gee_a$asthma
summary(mod_current)
```

Change from baseline, overall

```{r}
emmeans::emmeans(
  mod_current, 
  trt.vs.ctrl ~ data_schoolyear | student_race, 
  type = "response",
  weights = "proportional"
  )
```

YOY change, overall

```{r}
emmeans::contrast(emmeans::emmeans(
  mod_current, 
 trt.vs.ctrl ~ data_schoolyear | student_race, 
  type = "response",
  weights = "proportional"
  ), "consec")
```


```{r}
emmeans::contrast(emmeans::emmeans(
  mod_current, 
 trt.vs.ctrl ~ data_schoolyear | student_race, 
  type = "response",
  weights = "proportional"
  ), "poly")
```

Gender-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_gender | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Gender diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "pairwise"
)
```

Gender diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "del.eff"
)
```

Race/ethnicity-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_race | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Race/ethnicity diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "pairwise"
)
```

Race/ethnicity diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "del.eff"
)
```










# 504 Outcome

Model summary

```{r}
mod_current <- gee_a$x504
summary(mod_current)
```

Change from baseline, overall

```{r}
emmeans::emmeans(
  mod_current, 
  trt.vs.ctrl ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

YOY change, overall

```{r}
emmeans::contrast(emmeans::emmeans(
  mod_current, 
  ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  ), "consec")
```

Gender-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_gender | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Gender diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "pairwise"
)
```

Gender diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "del.eff"
)
```

Race/ethnicity-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_race | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Race/ethnicity diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "pairwise"
)
```

Race/ethnicity diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "del.eff"
)
```





















# IEP Outcome

Model summary

```{r}
mod_current <- gee_a$iep
summary(mod_current)
```

Change from baseline, overall

```{r}
emmeans::emmeans(
  mod_current, 
  trt.vs.ctrl ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

YOY change, overall

```{r}
emmeans::contrast(emmeans::emmeans(
  mod_current, 
  ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  ), "consec")
```

Gender-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_gender | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Gender diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "pairwise"
)
```

Gender diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "del.eff"
)
```

Race/ethnicity-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_race | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Race/ethnicity diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "pairwise"
)
```

Race/ethnicity diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "del.eff"
)
```




















# Attendance Outcome

Model summary

```{r}
mod_current <- gee_a$attendance
summary(mod_current)
```

Change from baseline, overall

```{r}
emmeans::emmeans(
  mod_current, 
  trt.vs.ctrl ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

YOY change, overall

```{r}
emmeans::contrast(emmeans::emmeans(
  mod_current, 
  ~ data_schoolyear, 
  type = "response",
  weights = "proportional"
  ), "consec")
```

Gender-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_gender | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Gender diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "pairwise"
)
```

Gender diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(
    mod_current, ~ student_gender, type = "response", weights = "proportional"
    ),
  "del.eff"
)
```

Race/ethnicity-specific estimates

```{r}
emmeans::emmeans(
  mod_current, 
  ~ student_race | data_schoolyear, 
  type = "response",
  weights = "proportional"
  )
```

Race/ethnicity diffs, pairwise

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "pairwise"
)
```

Race/ethnicity diffs, each vs. all others

```{r}
emmeans::contrast(
  emmeans::emmeans(mod_current, ~ student_race, type = "response",
  weights = "proportional"),
  "del.eff"
)
```

