---
title: "CPS Data Summary: Longitudinal Outcomes"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-06-09" 
  raw_data_cps:
    value:
      file: "UIC Asthma Status Data Albuterol SoW Shell_07252025.xlsx"
      sheet: "SY22, 23, 24 and 25 Asthma Stat"
      box_id: '1938297974039'
      clean_name: "student_counts"
  raw_data_schools_metadata:
    value: 
      file: "CPS Schools CCAs Regions.xlsx"
      sheet: NULL
      box_id: '2071252778107'
      clean_name: "schools"
  raw_data_utilization:
    # value:
    #   file: "CPS 2023-2025 utilization recode.xlsx"
    #   sheet: "Sheet1"
    #   box_id: '2017350498669'
    #   clean_name: "inhaler_events"
    value: # added 2026-03-25
      file: "CPS 2023-2025 utilization recode_3-23.xlsx"
      sheet: "Utilization Data"
      box_id: '2173811730030'
      clean_name: "inhaler_events"
      # force_preproc: TRUE
  raw_data_waves: # added 2026-01-21
    value:
      file: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
      box_id: "2109640911794"
      clean_name: "waves"
  raw_sy_dates: # added 2026-03-04
    value:
      box_id: "2154041346707"
      file: "schoolyear_dates.xlsx"
      sheet: "Sheet1"
      clean_name: "school_year_dates"
      # force_preproc: true
  enrollment_count_files: # box ids
    value: 
      sy_22: "2154049626576"
      sy_23: "2154041838698"
      sy_24: "2154043506963"
      sy_25: "2154043912572"
      sy_26: "2154047037077"
  force_preproc: FALSE
knit: |
  (function(input, ...) {
      rmarkdown::render(
        input, 
        output_file   = paste0("33753_cps_data_", Sys.Date()),
        output_dir    = file.path(dirname(input), "output"),
        envir         = globalenv()
      )
  })
---

```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}
# (maybe) create subdirectories for code, data, output
walk(nested_here(c("R", "data", "output", "output/tables")), ~{
  if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
  })

# load external functions
list.files(nested_here("R"), "fx.R$", full.names = TRUE) |> 
  walk(source)
```



# Data Processing

```{r}
# more precise schoolyear dates
data_dates <- full_join(
  read_rds(nested_here("data", "pre_school_year_dates.rds")) |> 
    filter(season == "fall") |> 
    distinct(data_schoolyear, start) |> 
    mutate(
      demog_date = start + duration(20, units = "days"),
      .keep      = "unused"
      ),
  
  read_rds(nested_here("data", "pre_school_year_dates.rds")) |> 
    filter(season == "spring") |> 
    distinct(data_schoolyear, end) |> 
    mutate(
      asthma_date     = end,
      attendance_date = end,
      .keep           = "unused"
      ),
  
  by = "data_schoolyear"
  )
```

```{r}
use_counts <- full_join(
  data_dates |> 
    mutate(admin_date = asthma_date),

  read_rds(nested_here("data", "mrg_inhaler_events.rds")) |> 
    select(school_id, admin_date),
  
  by = "admin_date"
) |> 
  
  arrange(admin_date) |> 
  fill(data_schoolyear, .direction = "up") |> 
  filter(!is.na(school_id)) |> 
  droplevels() |> 
  count(data_schoolyear, school_id, name = "n_inh_events")
```

```{r}
# read_rds(nested_here("data", "mrg_inhaler_events.rds")) |> 
#     select(school_id, admin_date)
# use_counts
```


```{r}
school_chars <- right_join(
  data_dates,
  read_rds(nested_here("data", "analytic_school_lvl.rds")),
  by = "data_schoolyear"
  ) |> 
  select(-matches("events")) |> 
  
  full_join(
    readxl::read_xls(
      nested_here("data", "CPS_EJI_EXCEL.xls"),
      na = c("", "-999")
      ) |> 
      janitor::clean_names() |>  
      distinct(school_id, rpl_ser) |> 
      mutate(school_id = as.character(school_id)),
    by = "school_id"
  ) |> 
  filter(!is.na(data_schoolyear))

school_chars <- school_chars |> 
  full_join(
    school_chars |> 
      select(data_schoolyear, school_id, matches("pct_student_race")) |> 
      pivot_longer(matches("pct_student_race")) |>
      filter(value == max(value), .by = c(data_schoolyear, school_id)) |> 
      mutate(
        re_profile = case_when(
          value >= 66.6 ~ gsub("pct_student_race", "maj", name),
          .default = "no_maj"
        ),
        .keep = "unused"
      ) 
  ) |> 
  
  full_join(
    school_chars |> 
      select(data_schoolyear, school_id, 
             student_count, count_student_gender_male,
             count_student_gender_unknown) |> 
      mutate(
        gender_profile = cut(
          count_student_gender_male / 
            (student_count - coalesce(count_student_gender_unknown, 0)),
          breaks         = c(0, .333, .666, 1),
          include.lowest = TRUE,
          labels         = c("majority female", "balanced", "majority male")
          ),
        .keep = "unused"
        )
  )
```


```{r}
ipw_data <- school_chars |> 
  
  rowwise() |> 
  mutate(
    sy_days         = 176,
    program_cum_days = max(
      as.numeric(difftime(asthma_date, impl_date, units = "days")), 
      0),
    
    program_sy_days = min(program_cum_days, sy_days)
    ) |> 
  ungroup() |> 
  
  mutate(
    sy_stu_days         = sy_days*student_count,
    program_sy_stu_days = program_sy_days*student_count,
    program_sy_prop     = program_sy_days / sy_days
  ) |> 
  
  left_join(use_counts, by = c("data_schoolyear", "school_id")) |> 
  mutate(
    n_events = coalesce(n_inh_events, 0),
    n_inh_rate = case_when(
      program_sy_stu_days == 0 ~ 0,
      .default = (n_events / program_sy_stu_days)*1000
      ),
    n_events_cumulative = cumsum(n_events),
    
    any_events = factor(I(n_events > 0)),
    any_events_cumulative = factor(I(n_events_cumulative > 0)),
    any_events_l3 = case_when(
      program_sy_days == 0 ~ "NA - Not Enrolled",
      .default = any_events
      ),
    any_events_cumulative_l3 = case_when(
      program_sy_days == 0 ~ "NA - Not Enrolled",
      .default = any_events_cumulative
      ),
    .by = school_id,
    .before = 1
    )

```



Q: school-level utilization sufficient? interested in demog-specific utilization?

Demographic data (year start)

```{r}
# schools overall, annual data
school_tby(
  y_var = "",
  x_var =  c(
            "notest(n_schools)", "grade_level", "notest(region)",
            "rpl_ser",
            "student_count", 
            "pct_student_race_american_indian",
            "pct_student_race_asian_hawaiian_pacific_islander",
            "pct_student_race_black_non_hispanic",
            "pct_student_race_hispanic", "pct_student_race_multi",
            "pct_student_race_unknown", "pct_student_race_white_non_hispanic",
            "notest(pct_student_race_middle_eastern_or_north_african)",
            "pct_student_gender_female", "pct_student_gender_male",
            "pct_student_gender_non_binary", "pct_student_gender_unknown",
            "pct_student_504_yes", "pct_iep_yes", "pct_esl_yes",
            "pct_frm_yes", "pct_homelessness_yes",
            "average_attendance",
            "pct_asthma_yes", "pct_food_allergy_yes",
            "pct_non_food_allergy_yes"
        ),
  df    = ipw_data
) |> 
  structure(
    title       = "School Characteristics by School Year (Day 20)",
    sheetname   = "descr_school_char_by_year",
    description = "Summary statistics of school-level characteristics and demographics. Variables from CPS are tabulated separately for each school year (2021-2022 to 2024-2025). Demographics are as of Day 20 of each school year. Chronic condition and attendance are as of the end of the school year."
  )
```



```{r}
ipw_data |> 
  relocate(matches("event"), matches("inh"))
```


```{r}
# schools overall, annual data
school_tby(
  y_var = "wave",
  x_var = c(
    # "wave",
    "as.factor(impl_date)",
    "factor(sy_days)", "factor(program_sy_days)", "factor(program_cum_days)",
    "n_events",
    "n_events_cumulative",
    "any_events",
    "any_events_cumulative",
    "any_events_l3",
    "any_events_cumulative_l3",
    "n_inh_rate"
    ), 
    df    = ipw_data |> 
    filter(!data_schoolyear %in% c("2021-2022", "2022-2023"))
    ) |> 
  rename(variable = 2) |> 
  
  filter(Total != "0 (0.0%)") |> 
  structure(
    title       = "Program Utilization by School Year",
    sheetname   = "school_util_by_year",
    description = "Summary statistics of school-level program enrollment and inhaler utilization. Data are as of the end of each school year."
  ) #|>
  
  # write_table("school_chars_by_sy")
```

Inhaler events per program days (cumulative)

```{r}
mod_null <- glm(
  n_events_cumulative ~ 1,
  offset = log(program_cum_days),
  family = poisson(),
  data   = ipw_data |> 
    filter(data_schoolyear == "2024-2025")
  ) 
summary(mod_null)
```
Predict inhaler events up to end of 2024-2025

```{r}
# covariates
ipw_data_2425 <- full_join(
  ipw_data |> 
    filter(data_schoolyear %in% c("2023-2024", "2024-2025")) |> 
    filter(!is.na(student_count)) |> 
    select(school_id, data_schoolyear, student_count, average_attendance, starts_with("pct"),
           matches("profile"),
           matches("days")) |> 
    arrange(data_schoolyear) |> 
    group_split(data_schoolyear) |> 
    reduce(
      full_join, 
      by     = c("school_id"),
      suffix = c(".23_24", ".24_25")
      ),
  
  # exposure vars
  ipw_data |> 
    filter(!is.na(student_count)) |> 
    filter(data_schoolyear == "2024-2025") |> 
    rename(exposure_date = asthma_date) |> 
    select(
      school_id, grade_level, rpl_ser, wave,
      exposure_date,
      n_events_cumulative, program_cum_days
    )
)
  
```

```{r}
ipw_numer <- glm(
   n_events_cumulative ~ 1,
   offset = log(program_cum_days),
   data = ipw_data_2425,
   family = poisson(link = "log")
   )
```


```{r}
library(mgcv)
ipw_gam <- gam(
  n_events_cumulative ~ 
    s(student_count.23_24, bs = "cr") +
    re_profile.24_25 +
    gen_profile.24_25 +
    s(average_attendance.23_24, bs = "cr") +
    s(pct_asthma_yes.23_24, bs = "cr") +
    s(student_count.24_25, bs = "cr") +
    s(average_attendance.24_25, bs = "cr") +
    s(pct_asthma_yes.24_25, bs = "cr") +
    s(pct_food_allergy_yes.24_25, bs = "cr") +
    s(pct_non_food_allergy_yes.24_25, bs = "cr") +
    s(pct_student_504_yes.24_25, bs = "cr") +
    s(pct_iep_yes.24_25, bs = "cr") +
    s(pct_esl_yes.24_25, bs = "cr") +
    s(pct_frm_yes.24_25, bs = "cr") +
    s(pct_homelessness_yes.24_25, bs = "cr") +
    s(rpl_ser, bs = "cr") + 
    grade_level + 
    wave,
  offset = log(program_cum_days),
  data   = ipw_data_2425,
  family = poisson(link = "log")
  )

summary(ipw_gam)
```



```{r}
stab_ipw <- 
  dpois(ipw_data_2425$n_events_cumulative, predict(ipw_numer, type = "response")) /
  dpois(ipw_data_2425$n_events_cumulative, predict(ipw_gam, type = "response"))
stab_ipw[which(stab_ipw < quantile(stab_ipw, .05))] <- quantile(stab_ipw, .05)
stab_ipw[which(stab_ipw > quantile(stab_ipw, .95))] <- quantile(stab_ipw, .95)
ipw_data_2425$stab_ipw <- stab_ipw
plot(stab_ipw)
```






```{r}
attd_mod <- gam(
  cbind(days_present, days_absent) ~ 
    s(util_rate, bs = "cr") + 
    s(student_count.23_24, bs = "cr") + 
    s(pct_asthma_yes.23_24, bs = "cr") + 
    s(student_count.24_25, bs = "cr") + 
    s(pct_asthma_yes.24_25, bs = "cr") + 
    s(pct_student_race_american_indian.24_25, bs = "cr") + 
    s(pct_student_race_asian_hawaiian_pacific_islander.24_25, bs = "cr") +
    s(pct_student_race_black_non_hispanic.24_25, bs = "cr") + 
    s(pct_student_race_hispanic.24_25, bs = "cr") + 
    s(pct_student_race_multi.24_25, bs = "cr") +
    s(pct_student_race_unknown.24_25, bs = "cr") + 
    s(pct_student_race_white_non_hispanic.24_25, bs = "cr") + 
    s(pct_student_race_middle_eastern_or_north_african.24_25, bs = "cr") +
    s(pct_student_gender_female.24_25, bs = "cr") + 
    s(pct_student_gender_male.24_25, bs = "cr") + 
    s(pct_student_gender_non_binary.24_25, bs = "cr") + 
    s(pct_student_gender_unknown.24_25, bs = "cr") +
    s(pct_food_allergy_yes.24_25, bs = "cr") + 
    s(pct_non_food_allergy_yes.24_25, bs = "cr") + 
    s(pct_student_504_yes.24_25, bs = "cr") + 
    s(pct_iep_yes.24_25, bs = "cr") + 
    s(pct_esl_yes.24_25, bs = "cr") + 
    s(pct_frm_yes.24_25, bs = "cr") + 
    s(pct_homelessness_yes.24_25, bs = "cr") + 
    s(rpl_ser, bs = "cr") + 
    grade_level + 
    wave,
  data = ipw_data_2425,
  family = binomial(link = "logit"),
  weights = stab_ipw
  )

summary(attd_mod)
```
























```{r}
library(splines)
ipw_denom <- glm(
  n_events_cumulative ~ 
    bs(student_count.23_24, knots = median(student_count.23_24)) + 
    bs(average_attendance.23_24, knots = median(average_attendance.23_24)) + 
    bs(pct_asthma_yes.23_24, knots = median(pct_asthma_yes.23_24)) + 
    bs(student_count.24_25, knots = median(student_count.24_25)) + 
    bs(average_attendance.24_25, knots = median(average_attendance.24_25)) + 
    bs(pct_asthma_yes.24_25, knots = median(pct_asthma_yes.24_25)) + 
    bs(pct_student_race_american_indian.24_25, knots = median(pct_student_race_american_indian.24_25)) + 
    bs(pct_student_race_asian_hawaiian_pacific_islander.24_25, knots = median(pct_student_race_asian_hawaiian_pacific_islander.24_25)) +
    bs(pct_student_race_black_non_hispanic.24_25, knots = median(pct_student_race_black_non_hispanic.24_25)) + 
    bs(pct_student_race_hispanic.24_25, knots = median(pct_student_race_hispanic.24_25)) + 
    bs(pct_student_race_multi.24_25, knots = median(pct_student_race_multi.24_25)) +
    # bs(pct_student_race_unknown.24_25, knots = median(pct_student_race_unknown.24_25)) + 
    bs(pct_student_race_white_non_hispanic.24_25, knots = median(pct_student_race_white_non_hispanic.24_25)) + 
    # bs(knots = median(# ))pct_student_race_middle_eastern_or_north_african.24_25 + 
    bs(pct_student_gender_female.24_25, knots = median(pct_student_gender_female.24_25)) + 
    bs(pct_student_gender_male.24_25, knots = median(pct_student_gender_male.24_25)) + 
    # bs(pct_student_gender_non_binary.24_25, knots = median(pct_student_gender_non_binary.24_25)) + 
    # bs(knots = median(# ))pct_student_gender_unknown.24_25 + 
    bs(pct_food_allergy_yes.24_25, knots = median(pct_food_allergy_yes.24_25)) + 
    # bs(pct_non_food_allergy_yes.24_25, knots = median(pct_non_food_allergy_yes.24_25)) + 
    bs(pct_student_504_yes.24_25, knots = median(pct_student_504_yes.24_25)) + 
    bs(pct_iep_yes.24_25, knots = median(pct_iep_yes.24_25)) + 
    bs(pct_esl_yes.24_25, knots = median(pct_esl_yes.24_25)) + 
    bs(pct_frm_yes.24_25, knots = median(pct_frm_yes.24_25)) + 
    bs(pct_homelessness_yes.24_25, knots = median(pct_homelessness_yes.24_25)) + 
    bs(rpl_ser) + 
    grade_level + 
    wave,
  offset = log(program_cum_days),
  data   = ipw_data_2425,
  family = poisson(link = "log")
  )

summary(ipw_denom)
```


```{r}
stab_ipw <- 
  dpois(ipw_data_2425$n_events_cumulative, predict(ipw_numer, type = "response")) /
  dpois(ipw_data_2425$n_events_cumulative, predict(ipw_denom, type = "response"))
stab_ipw[which(stab_ipw < quantile(stab_ipw, .01))] <- quantile(stab_ipw, .01)
stab_ipw[which(stab_ipw > quantile(stab_ipw, .99))] <- quantile(stab_ipw, .99)
ipw_data_2425$stab_ipw <- stab_ipw
plot(stab_ipw)
```

```{r}
# outcomes
ipw_data_2425 <- ipw_data_2425 |> 
  mutate(
    util_rate    = n_events_cumulative / program_cum_days,
    days_present = round((average_attendance.24_25/100)*student_count.24_25),
    days_absent  = round((1 - average_attendance.24_25/100)*student_count.24_25),
    .before = 1
    )
```


```{r}

attd_mod <- glm(
  cbind(days_present, days_absent) ~ 
    util_rate + 
    student_count.23_24 + pct_asthma_yes.23_24 + 
    student_count.24_25 + pct_asthma_yes.24_25 + 
    pct_student_race_american_indian.24_25 + pct_student_race_asian_hawaiian_pacific_islander.24_25 +
    pct_student_race_black_non_hispanic.24_25 + pct_student_race_hispanic.24_25 + pct_student_race_multi.24_25 +
    pct_student_race_unknown.24_25 + pct_student_race_white_non_hispanic.24_25 + 
    # pct_student_race_middle_eastern_or_north_african.24_25 + 
    pct_student_gender_female.24_25 + 
    pct_student_gender_male.24_25 + 
    pct_student_gender_non_binary.24_25 + 
    # pct_student_gender_unknown.24_25 + 
    pct_food_allergy_yes.24_25 + 
    pct_non_food_allergy_yes.24_25 + 
    pct_student_504_yes.24_25 + 
    pct_iep_yes.24_25 + 
    pct_esl_yes.24_25 + 
    pct_frm_yes.24_25 + 
    pct_homelessness_yes.24_25 + 
    grade_level + 
    rpl_ser + 
    wave,
  data = ipw_data_2425,
  family = binomial(link = "logit"),
  weights = stab_ipw
  )

summary(attd_mod)
```





```{r}
# test changes from year to year (all pairwise)
read_rds(nested_here("data", "analytic_school_lvl.rds")) |> 
  filter(!is.na(student_count)) |> 
  droplevels() |>
  pivot_longer(
    c(student_count, average_attendance, matches("pct")),
    names_to = "x"
  ) |> 
  
  pivot_wider(
    names_from = data_schoolyear,
    id_cols    = c(school_id, x)
  ) |> 
  
  mutate(
    `(2022-2023) - (2021-2022)` = `2022-2023` - `2021-2022`,
    `(2023-2024) - (2021-2022)` = `2023-2024` - `2021-2022`,
    `(2023-2024) - (2022-2023)` = `2023-2024` - `2022-2023`,
    `(2024-2025) - (2021-2022)` = `2024-2025` - `2021-2022`,
    `(2024-2025) - (2022-2023)` = `2024-2025` - `2022-2023`,
    `(2024-2025) - (2023-2024)` = `2024-2025` - `2023-2024`
  ) |> 
  
  select(school_id, x, matches("\\(")) |> 
  pivot_longer(where(is.numeric), names_to = "contrast", values_to = "abs_change") |> 
  
  summarise(
    n_schools = sum(!is.na(abs_change)),
    broom::tidy(wilcox.test(abs_change, exact = FALSE, correct = TRUE, conf.int = TRUE)),
    .by       = c(x, contrast)
  ) |> 
  rename(med.change = estimate, med.change.conf.low = conf.low, med.change.conf.high = conf.high) |> 
  mutate(
    p.adjust        = p.adjust(p.value, "holm"), 
    p.adjust.n      = sum(!is.na(p.value)),
    p.adjust.method = "holm",
    .after          = p.value,
    .by             = x
  )  |> 
  
  structure(
    title       = "Test Change in School Characteristics Over Time",
    sheetname   = "test_school_char_change",
    description = "Annual school-level numeric variables (e.g., attendance, % with asthma) were tested for evidence of change over time. For each variable, within-school change was calculated from schoolyear to schoolyear (all pairwise schoolyear comparisons). School-level change between each pair of years was then tested for symmetry around zero (i.e., no change over time) by a one-sample Wilcoxon signed-rank test. p-values for the set of 6 pairwise year comparisons for each variable were adjusted by the Holm method."
  ) |>
  
  write_table("test_school_chars_by_sy.rds")
```


```{r}
# Describe School Chars (2023-2024) by Inhaler Events ---------------------

# any events, 2023-2024 only, separating late implementers
school_tby(
  y_var = "any_events_l3",
  df    = read_rds(nested_here("data", "analytic_school_lvl.rds")) |> 
    filter(data_schoolyear %in% c("2023-2024")),
  tby_ctrl = arsenal::tableby.control(
    test           = FALSE,
    digits.n	      = NA,
    numeric.stats  = c("medianrange", "q1q3",  "meansd")
  ),
) |> 
  
  structure(
    title       = "School Characteristics by Leven of Inhaler Utilization, SY 2023-2024",
    sheetname   = "descr_util_groups_23_24",
    description = "Summary statistics of school-level characteristics and demographics in the 2023-2024 schoolyear. Schools are grouped by inhaler utilization level in the 2023-2024 school year (0 events vs. 1+ events vs. N/A).") |> 
  
  write_table("school_chars_by_util.rds")


# test assn between school chars and school-level utilization
full_join(
  
  c("grade_level", "region_3grp",  "region_2grp" ) |> 
    set_names() |> 
    map_dfr(~{
      test_categ(.x, groups = "any_events_l3", data = read_df_sy("2023-2024"))
    }, .id = "x"),
  
  c("student_count", "average_attendance", 
    grep("pct", names(read_df_sy("2023-2024")), value = TRUE)) |> 
    set_names() |>  
    map_dfr(~{
      test_ordered(.x, groups = "any_events_l3", data = read_df_sy("2023-2024"))
    }, .id = "x")
  
) |> 
  mutate(data_schoolyear = "2023-2024", .before = 1) |> 
  relocate(data_schoolyear, x, x_pair, group1, group2, n, p.value, matches("p.adj")) |> 
  
  structure(
    title       = "Test Differences in School Characteristics by Level of Inhaler Utilization, SY 2023-2024",
    sheetname   = "test_util_group_diffs_23_24",
    description = "Three-group and pairwise comparisons of school characteristics by inhaler utilization level in the SY 2023-2024 schoolyear. Numeric variables are compared across groups (0 events, 1+ events, NA) by Kruskal-Wallis tests (all 3 groups) and pairwise Wilcoxon rank-sum tests (each pair of 2 groups). Categorical variables (e.g., grade level) are compared across groups (0 events, 1+ events, NA) Fisher's exact tests (all 3 groups and each pair of 2 groups). p-values for the set of three pairwise comparisons for each variable were adjusted by the Holm method.") |> 
  
  write_table("test_school_chars_by_util.rds")

```





```{r}
ipw::ipwtm(
  exposure = n_inh_rate,
  family = "ordinal",
  link = "logit",
  numerator = ~ 1,
  denominator = ~ data_schoolyear + program_sy_days + grade_level + wave +
    splines::ns(pct_student_race_american_indian) +
    splines::ns(pct_student_race_asian_hawaiian_pacific_islander),
  id = school_id,
  data = ipw_data,
  timevar = as.numeric(data_schoolyear),
  type = "all"
  )
```



after https://www.andrewheiss.com/blog/2020/12/01/ipw-binary-continuous/

(cross-sectional) iptw: 

numer_mod = glm(util ~ 1, data = data)
numer = dnorm(data$util, predict(numer_nod), sd = numer_mod$residuals)

denom_mod = glm(util ~ demog + asthma, data = data)
denom = dnorm(data$util, predict(denom_nod), sd = denom_mod$residuals)

data$ipw = numer / denom
attendance_mod = glm(attendance ~ util, data = data, weights = ipw)

alt: use ipw::ipwpoint()






```{r}

```














# Describe School Chars by Year

```{r, eval = FALSE}
# source(nested_here("R", "DESCRIBE_SCHOOLS.R"))
```

Results preview:

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_SCHOOLS.R")) |> 
  read_table() |> 
  map(head, 10) |> 
  map(knitr::kable, digits = 3) |> 
  walk(print)
```


# Utilization Detail

```{r}
tableby_with <- function(df, ...) {
  arsenal::tableby(
    ...,
    data = df
    )
}
```

```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_UTILIZATION.R"))
```

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_UTILIZATION.R")) |> 
  read_table() 
```




What predicts utilization? Poisson (or similar) for inhaler uses, 
with a student-time (in program) offset. 
Adjust for school-level chars. 

- [ ] tabulate uses per school per year

```{r}

```



```{r}
util_ct_w_offset <- left_join(
  read_rds(nested_here("data", "analytic_school_lvl.rds")),
  read_rds(nested_here("data", "mrg_inhaler_events.rds")) |> 
    count(school_id, data_schoolyear, name = "n_incidents"),
  by = c("school_id", "data_schoolyear")
  ) |> 
  left_join(
    read_rds(nested_here("data", "pre_school_year_dates.rds")) |> 
      filter(season != "summer") |> 
      summarise(
        data_date = max(end),
        sy_startdate = min(start),
        sy_enddate = data_date,
        .by = data_schoolyear
        ),
    by = "data_schoolyear"
    ) |> 
  filter(!is.na(student_count)) |> 
  relocate(matches("date")) |> 
  rowwise() |> 
  mutate(
    n_incidents        = coalesce(n_incidents, 0),
    sy_days = as.numeric(difftime(sy_enddate, sy_startdate, units = "days")),
    impl_days_cum = max(
      0,
      as.numeric(difftime(data_date, impl_date, units = "days"))
      ),
    impl_days_sy = min(sy_days, impl_days_cum),
    
    studentdays_sy         = sy_days*student_count,
    impl_studentdays_sy    = impl_days_sy*student_count,
    n_inc_per_1k_impl_studays = (n_incidents / impl_studentdays_sy)*1000,
    n_inc_per_1k_studays      = (n_incidents / studentdays_sy)*1000,
    .before                = 1
    ) |> 
  ungroup() 
  
util_ct_w_offset |> arrange(desc(n_inc_per_1k_studays))
```






- [ ] calculate student-program-time per school per year 
      (time btw program implementation and end of school year)
- [ ] calculate use rate per school per year
- [ ] plot 
    - [ ] scatter: rate over time by impl date
- [ ] model uses (count with student-time offset) 
      as predicted by school demographics,
      school characteristics, calendar time (school year), time in program

Categorize schools by rate of utilization (quartiles). Categorize 
schools by outcome rates (quartiles). Tabulate, plot by year. 


# Student Outcomes


Describe trends in student outcomes (student group level) 
over time (asthma, 504 plans, ieps, attendance).
Allow for demog-by-time interaction terms.


# Outcomes x Utilization


Latent class analysis/latent transition analysis. Categorize by low/high utilizers 
and low/high need. 

Describe trends in outcomes as predicted by time-in-program, utilization 
(treat as time-varying exposure?). Causal analysis w 
time-varying covariates (confounders): use IPTW for utilization?





after https://www.andrewheiss.com/blog/2020/12/01/ipw-binary-continuous/

(cross-sectional) iptw: 

numer_mod = glm(util ~ 1, data = data)
numer = dnorm(data$util, predict(numer_nod), sd = numer_mod$residuals)

denom_mod = glm(util ~ demog + asthma, data = data)
denom = dnorm(data$util, predict(denom_nod), sd = denom_mod$residuals)

data$ipw = numer / denom
attendance_mod = glm(attendance ~ util, data = data, weights = ipw)

alt: use ipw::ipwpoint()

# Write Results

```{r}
out_tables <- list(
  About = t(data.frame(
    Date   = Sys.Date(),
    Author = "Rachel Lane",
    Title  = "SOW 2 Tables"
  )) |> 
  as.data.frame() |> 
  set_names("value") |> 
  rownames_to_column("name")
  )
out_tables[["About"]] <- rbind(
  out_tables[["About"]],
  params |> 
    keep_at(~ grepl("raw", .x)) |> 
    set_names_from_child("clean_name") |> 
    map_dfr(~data.frame(
      name = "Data Sources",
      value = glue::glue("File: {.x$file}. Sheet: {coalesce(.x$sheet, '1')}. Box file ID: {.x$box_id}")
      )))

```

```{r}
for (f in list.files(
  nested_here("output", "tables"),
  "sow3"
  )) {
  d <- read_rds(nested_here("output", "tables", f))
  out_tables[[attr(d, "sheetname")]] <- d
  out_tables$About <- out_tables$About |> 
    add_case(
      name = "Contents",
      value = glue::glue('Sheet: {attr(d, "sheetname")}. Title: {attr(d, "title")}. Description: {attr(d, "description")}')
    )
  }

```

```{r}
openxlsx::write.xlsx(
  out_tables,
  nested_here("output", paste0("sow3_tables_", Sys.Date(), ".xlsx")),
  asTable = TRUE
)
```


