176 student-days -- use as baseline


modify in descriptive tables

comments -- randomized trials in survey design








## What predicts utilization?


```{r}
attd_change <- school_change |> 
  filter(grepl("attendance", variable)) |> 
  pivot_wider(names_from = "variable", names_prefix = "chg_") |> 
  mutate(
    data_schoolyear = gsub("([0-9]+)\\.([0-9]+)-.*", "20\\1-20\\2", years),
    .keep = "unused"
    )
```


```{r}
library(tidyverse)

schools <- read_rds(here::here(
  "a12_cps_data/2026-07-22/data/schools_analytic.rds" 
  ))



schools_mod <- schools |> 
  filter(!is.na(rpl_ser)) |> 
  mutate(n_schools = "Overall", .before = 1) |> 
  mutate(across(matches("pct"), ~ .x*100)) |> 
  filter(!(data_schoolyear %in% c("2021-2022", "2022-2023"))) |> 
  mutate(data_schoolyear = factor(data_schoolyear, ordered = FALSE)) |> 
  droplevels() |> 
  
  left_join(attd_change, by = c("school_id", "data_schoolyear"))
```


```{r}
library(arsenal)
tableby(any_events ~ notest(n_schools) 
        + grade_level 
        + notest(region) 
        + wave 
        + student_count 
        + re_majority 
        + pct_male 
        + pct_female 
        + pct_nonbinary 
        + rpl_ser 
        + asthma_pct 
        + food_allergy_pct 
        + non_food_allergy_pct 
        + student_504_pct 
        + iep_pct 
        + esl_pct 
        + frm_pct 
        + homelessness_pct
        + attendance + chg_attendance
        + attendance_asthma + chg_attendanceasthma,
        strata = data_schoolyear,
        data = schools_mod,
        control = tableby.control(
          total          = FALSE,
          digits.n	     = NA,
          digits        = 2,
          numeric.test = "kwt",
          ordered.test = "kwt",
          numeric.stats  = c("medianrange", "q1q3",  "meansd")
          )
        ) |> 
  summary(text = TRUE) |> 
  as.data.frame() |> 
  
  openxlsx::write.xlsx(
    nested_here("output", "tables", glue::glue("{Sys.Date()}_t2_event_predictors_univar.xlsx")),
    asTable = TRUE
    )
```


```{r}
schools_mod |> 
  pivot_longer(starts_with("chg_att")) |> 
  ggplot(aes(value, color = grade_level)) +
  geom_density() +
  facet_grid(name ~ data_schoolyear) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme(legend.position = "bottom")
```

```{r}
schools_mod |> 
  pivot_longer(starts_with("chg_att")) |> 
  ggplot(aes(value, color = re_majority)) +
  geom_density() +
  facet_grid(name ~ data_schoolyear) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "lightgray") +
  theme(legend.position = "bottom")
```


```{r}
schools_mod |> 
  pivot_longer(starts_with("chg_att")) |> 
  ggplot(aes(grade_level, value)) +
  geom_boxplot(width = 0.5, alpha = 0.75) +
  facet_grid(. ~ name) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.25)
```



more meaningful to report student-days gained or lost


```{r}
library(arsenal)
tableby(cut(chg_attendance, quantile(chg_attendance)) ~ notest(n_schools) 
        + grade_level 
        + notest(region) 
        + wave 
        + student_count 
        + re_majority 
        + pct_male 
        + pct_female 
        + pct_nonbinary 
        + rpl_ser 
        + asthma_pct 
        + food_allergy_pct 
        + non_food_allergy_pct 
        + student_504_pct 
        + iep_pct 
        + esl_pct 
        + frm_pct 
        + homelessness_pct
        + any_events + n_events,
        strata = data_schoolyear,
        data = schools_mod,
        control = tableby.control(
          total          = FALSE,
          digits.n	     = NA,
          digits        = 2,
          numeric.test = "kwt",
          ordered.test = "kwt",
          numeric.stats  = c("medianrange", "q1q3",  "meansd")
          )
        ) |> 
  summary(text = TRUE) |> 
  as.data.frame() |> 
  
  openxlsx::write.xlsx(
    nested_here("output", "tables", glue::glue("{Sys.Date()}_t2_attd_chg_pred_univar.xlsx")),
    asTable = TRUE
    )
```

```{r}
library(arsenal)
tableby(cut(chg_attendanceasthma, quantile(chg_attendanceasthma, na.rm = TRUE)) ~ notest(n_schools) 
        + grade_level 
        + notest(region) 
        + wave 
        + student_count 
        + re_majority 
        + pct_male 
        + pct_female 
        + pct_nonbinary 
        + rpl_ser 
        + asthma_pct 
        + food_allergy_pct 
        + non_food_allergy_pct 
        + student_504_pct 
        + iep_pct 
        + esl_pct 
        + frm_pct 
        + homelessness_pct
        + any_events + n_events,
        strata = data_schoolyear,
        data = schools_mod,
        control = tableby.control(
          total          = FALSE,
          digits.n	     = NA,
          digits        = 2,
          numeric.test = "kwt",
          ordered.test = "kwt",
          numeric.stats  = c("medianrange", "q1q3",  "meansd")
          )
        ) |> 
  summary(text = TRUE) |> 
  as.data.frame() |> 
  
  openxlsx::write.xlsx(
    nested_here("output", "tables", glue::glue("{Sys.Date()}_t2_asthmaattd_chg_pred_univar.xlsx")),
    asTable = TRUE
    )
```




## Step 1: Weights

"The treatment model can be defined as an association
model that "predicts" use of rescue medication as
some function of confounders." [@Mortimer2005]

First, calculate stabilized IPTW: "(marginal probability/conditional probability) or P(A)/P(A|W)" [@Mortimer2005]

Numerator: P(A) >> probability of exposure level
Denominator: P(A|W) >> probability of exposure level given confounders

P(A)

```{r}

```

```{r}
summary(m_iptw_numer <- geepack::geeglm(
  any_events_binary ~ data_schoolyear,
  family = binomial(link = "logit"),
  id     = factor(school_id),
  waves  = data_schoolyear,
  data   = schools_mod
))
```



```{r}
summary(m_iptw_denom <- geepack::geeglm(
  any_events_binary ~ data_schoolyear
    + poly(student_count, 3)
    + re_majority
    + grade_level
    + wave
    + poly(pct_male, 3)
    + poly(rpl_ser, 3)
    + poly(impl_days, 3)
    + poly(asthma_pct, 3)
    + poly(asthma_pct.lastyear, 3)
    + poly(student_504_pct, 3)
    + poly(iep_pct, 3)
    + poly(esl_pct, 3)
    + poly(frm_pct, 3)
    + poly(homelessness_pct, 3)
  ,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))

```

```{r}
# look for non-linear terms
# library(mgcv)
# summary(gam_mod <- gam(
#   any_events_binary ~ data_schoolyear
#     + s(student_count)
#     + re_majority + grade_level + wave
#     + s(pct_male)
#     + s(rpl_ser)
#     + s(impl_days, k = 8)
#     + s(asthma_pct)
#     + s(asthma_pct.lastyear)
#     + s(student_504_pct)
#     + s(iep_pct)
#     + s(esl_pct)
#     + s(frm_pct)
#     + s(homelessness_pct),
#   family = binomial(link = "logit"),
#   data   = schools_mod
# ))
```
```{r}
# summary(m_iptw_denom <- geepack::geeglm(
#   any_events_binary ~ data_schoolyear 
#     + student_count  
#     + re_majority + grade_level + wave
#     + poly(pct_male, 2) 
#     + rpl_ser
#     + impl_days
#     + asthma_pct
#     + asthma_pct.lastyear
#     + poly(student_504_pct, 2) 
#     + poly(iep_pct, 3) 
#     + esl_pct
#     + frm_pct
#     + poly(homelessness_pct, 2)
#   ,
#   family = binomial(link = "logit"),
#   id     = school_id,
#   waves  = data_schoolyear,
#   data   = schools_mod
# ))
```

```{r}
schools_mod <- schools_mod |> 
  mutate(
    stab_iptw = case_when(
      any_events_binary == 1 ~ (
        predict(m_iptw_numer, type = "response") /
          predict(m_iptw_denom, type = "response")
        ),
      any_events_binary == 0 ~ (
        (1 - predict(m_iptw_numer, type = "response")) /
          (1 - predict(m_iptw_denom, type = "response"))
        )
      )
    )

```

Step 2: Attendance

```{r}
schools_mod <- schools_mod |> 
  mutate(
    days_present  = round((student_count*176)*(attendance / 100)),
    days_absent   = round((student_count*176)*(1 - (attendance / 100))),
    days_expected = round((student_count*176)),
    .before       = 1
  )
```

```{r}
summary(mgcv::gam(
  cbind(days_present, days_absent) ~ 
    s(stab_iptw) + data_schoolyear*any_events_binary
    + s(student_count)
    + re_majority + grade_level + wave
    + s(pct_male)
    + s(rpl_ser)
    + s(impl_days, k = 8)
    + s(asthma_pct)
    + s(asthma_pct.lastyear)
    + s(student_504_pct)
    + s(iep_pct)
    + s(esl_pct)
    + s(frm_pct)
    + s(homelessness_pct)
  ,
  family = binomial(link = "logit"),
  # id     = school_id,
  # waves  = data_schoolyear,
  data   = schools_mod
))
```


```{r}
summary(geepack::geeglm(
  cbind(days_present, days_absent) ~ data_schoolyear 
    + s(stab_iptw, 3) 
    + any_events_binary
    
    + student_count
    + re_majority + grade_level + wave
    + s(pct_male, 2)
    + rpl_ser
    + impl_days
    + asthma_pct
    + asthma_pct.lastyear
    + s(student_504_pct, 2)
    + s(iep_pct, 3)
    + esl_pct
    + frm_pct
    + s(homelessness_pct, 2)
  ,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))
```















