## Step 1: Weights

"The treatment model can be defined as an association
model that ‘‘predicts’’ use of rescue medication as
some function of confounders." [@Mortimer2005]

First, calculate stabilized IPTW: "(marginal probability/conditional probability) or P(A)/P(A|W)" [@Mortimer2005]

Numerator: P(A) >> probability of exposure level
Denominator: P(A|W) >> probability of exposure level given confounders

P(A)

```{r}
library(tidyverse)

schools <- read_rds(here::here(
  "a12_cps_data/2026-07-22/data/schools_analytic.rds" 
  ))



schools_mod <- schools |> 
  filter(!(data_schoolyear %in% c("2021-2022", "2022-2023"))) |> 
  droplevels()
```

```{r}
writeLines(names(schools_mod), sep = " + ", con = "clipboard")
```

```{r}
summary(schools_mod)
```

```{r}
(m_iptw_numer <- geepack::geeglm(
  any_events_binary ~ ordered(data_schoolyear),
  family = binomial(link = "logit"),
  id     = factor(school_id),
  waves  = as.numeric(ordered(data_schoolyear)),
  data   = schools_mod
))
```

```{r}
summary(schools_mod$impl_days)
```


```{r}
library(arsenal)
tableby(any_events ~ student_count +
          grade_level + region + wave + 
          rpl_ser + 
          re_majority + pct_male +
          asthma_pct + food_allergy_pct + non_food_allergy_pct + 
          student_504_pct + iep_pct + 
          esl_pct + frm_pct + homelessness_pct + 
          attendance + attendance_asthma,
        strata = data_schoolyear,
        data = schools_mod) |> 
  summary()

```

```{r}
summary(m_iptw_numer <- geepack::geeglm(
  any_events_binary ~ data_schoolyear,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))
```


```{r}
summary(m_iptw_denom <- geepack::geeglm(
  any_events_binary ~ data_schoolyear 
    + poly(student_count, 3)  
    + re_majority + grade_level + wave
    + poly(pct_male, 3) 
    + poly(rpl_ser, 3)  
    + poly(impl_days, 3)
    + poly(asthma_pct, 3) 
    + poly(asthma_pct.lastyear, 3)
    + poly(student_504_pct, 3) 
    + poly(iep_pct, 3) 
    + poly(esl_pct, 3) 
    + poly(frm_pct, 3) 
    + poly(homelessness_pct, 3)
  ,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))

```

```{r}
# look for non-linear terms
# library(mgcv)
# summary(gam_mod <- gam(
#   any_events_binary ~ data_schoolyear 
#     + s(student_count) 
#     + re_majority + grade_level + wave
#     + s(pct_male) 
#     + s(rpl_ser)
#     + s(impl_days, k = 8)
#     + s(asthma_pct) 
#     + s(asthma_pct.lastyear)
#     + s(student_504_pct) 
#     + s(iep_pct) 
#     + s(esl_pct)
#     + s(frm_pct)
#     + s(homelessness_pct),
#   family = binomial(link = "logit"),
#   data   = schools_mod
# ))
```
```{r}
summary(m_iptw_denom <- geepack::geeglm(
  any_events_binary ~ data_schoolyear 
    + student_count  
    + re_majority + grade_level + wave
    + poly(pct_male, 2) 
    + rpl_ser
    + impl_days
    + asthma_pct
    + asthma_pct.lastyear
    + poly(student_504_pct, 2) 
    + poly(iep_pct, 3) 
    + esl_pct
    + frm_pct
    + poly(homelessness_pct, 2)
  ,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))
```

```{r}
schools_mod <- schools_mod |> 
  mutate(
    stab_iptw = case_when(
      any_events_binary == 1 ~ (
        predict(m_iptw_numer, type = "response") /
          predict(m_iptw_denom, type = "response")
        ),
      any_events_binary == 0 ~ (
        (1 - predict(m_iptw_numer, type = "response")) /
          (1 - predict(m_iptw_denom, type = "response"))
        )
      )
    )

```

Step 2: Attendance

```{r}
schools_mod <- schools_mod |> 
  mutate(
    days_present  = round((student_count*176)*(attendance / 100)),
    days_absent   = round((student_count*176)*(1 - (attendance / 100))),
    days_expected = round((student_count*176)),
    .before       = 1
  )
```

```{r}
summary(geepack::geeglm(
  cbind(days_present, days_absent) ~ data_schoolyear 
    + poly(stab_iptw, 3) 
    + any_events_binary
    
    + student_count
    + re_majority + grade_level + wave
    + poly(pct_male, 2)
    + rpl_ser
    + impl_days
    + asthma_pct
    + asthma_pct.lastyear
    + poly(student_504_pct, 2)
    + poly(iep_pct, 3)
    + esl_pct
    + frm_pct
    + poly(homelessness_pct, 2)
  ,
  family = binomial(link = "logit"),
  id     = school_id,
  waves  = data_schoolyear,
  data   = schools_mod
))
```

