---
title: "CPS Data Summary: Longitudinal Outcomes"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-07-28"
  school_profile_api_endpoints: 
    value: 
      sy_2122: "2dem-8rq7"
      sy_2223: "9a5f-2r4p"
      sy_2324: "cu4u-b4d9"
      sy_2425: "3dhs-m3w4"
  raw_data_cps_counts:
    value:
      box_id: "2332953277707"
      filename: "CPS SOW Data Pull 7-7-2026 Updated Gender.xlsx"
      sheet: "Updated SY22, 23, 24 and 25 Ast"
  raw_data_utilization:
    value: 
      box_id: '2173811730030'
      filename: "CPS 2023-2025 utilization recode_3-23.xlsx"
      sheet: "Utilization Data"
  raw_data_waves:
    value:
      box_id: "2109640911794"
      filename: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
  raw_data_sy_dates:
    value:
      box_id: "2154041346707"
      filename: "schoolyear_dates.xlsx"
      sheet: "Sheet1"
# knit: |
#   (function(input, ...) {
#       rmarkdown::render(
#         input, 
#         output_file   = paste0("33753_cps_data_", Sys.Date()),
#         output_dir    = file.path(dirname(input), "output"),
#         envir         = globalenv()
#       )
#   })
---

```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}
# # (maybe) create subdirectories for code, data, output
# walk(nested_here(c("R", "data", "output", "output/tables")), ~{
#   if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
#   })
```

```{r}
# preproc
# source(nested_here("R", "PREPROC.R"))
```

```{r}
# fetch school characteristics from chicago data portal
school_chars <- params$school_profile_api_endpoints |> 
  map_dfr(~{
    glue::glue("https://data.cityofchicago.org/api/v3/views/{.x}/query.json") |> 
      httr::GET() |> 
      pluck("content") |> 
      rawToChar() |> 
      jsonlite::fromJSON() |> 
      janitor::clean_names() |> 
      select(school_id, short_name, primary_category, matches("(lat)|(long)|(coord)"))
  }) |> 
  mutate(
    school_match_name = gsub("_", "", snakecase::to_screaming_snake_case(short_name)),
    grade_level       = fct_collapse(
      primary_category,
      "Elementary/Middle School" = c("ES", "MS"),
      "High School"              = c("HS")
      )
    ) |> 
  distinct()
```

```{r}
# fetch city region boundaries from chicago data portal; merge to schools by location
school_chars <- httr2::request("https://data.cityofchicago.org/api/v3/views/spyv-p8fk/query.json") |> 
  httr2::req_perform() |> 
  httr2::resp_body_json() |>
  enframe() |> 
  mutate(
    region_name = unlist(map(value, ~ pluck(.x, "region_nam"))),
    geom        = map(value, ~{
      .x |> 
        pluck("the_geom") |>
        pluck("coordinates") |> 
        unlist() |> 
        matrix(byrow = TRUE, ncol = 2) |> 
        as.data.frame() |> 
        sf::st_as_sf(coords = c("V1", "V2")) |> 
        sf::st_combine() |>  
        sf::st_cast("POLYGON")
      }),
    
    schools = map(geom, ~{
      sf::st_intersection(
        school_chars |> 
          sf::st_as_sf(coords = c("school_longitude", "school_latitude")),
        .x
        ) |> 
        sf::st_drop_geometry()
      })
  ) |> 
  select(schools, region_name) |> 
  unnest(schools)


```





# Describe School Chars by Year

```{r}
schools <- read_rds(nested_here("data/schools_analytic.rds")) 
```

```{r}
library(arsenal)
tableby(
  data_schoolyear ~ n_schools 
  + grade_level 
  + region 
  + wave 
  + student_count 
  + re_majority 
  + pct_male 
  + pct_female 
  + pct_nonbinary 
  + rpl_ser 
  + asthma_pct 
  + food_allergy_pct 
  + non_food_allergy_pct 
  + student_504_pct 
  + iep_pct 
  + esl_pct 
  + frm_pct 
  + homelessness_pct
  + attendance 
  + days_per_stu_expected
  + days_per_stu_present
  + days_per_stu_absent
  + attendance_asthma 
  + days_per_stu_present_asthma
  + days_per_stu_absent_asthma
  + n_events
  + any_events 
  ,
  data    = schools,
  control = tableby.control(
    test           = FALSE,
    total          = FALSE,
    digits.n	     = NA,
    digits        = 2,
    numeric.stats  = c("Nmiss", "medianrange", "q1q3",  "meansd")
    )
  ) |> 
  summary(text = TRUE) |> 
  as.data.frame() |>
  write_rds(nested_here("output", "t1_descriptive.rds"))

read_rds(nested_here("output", "t1_descriptive.rds"))
```


```{r}
read_rds(nested_here("output", "t1_descriptive.rds")) |> 
  openxlsx::write.xlsx(
    nested_here("output", "tables", "t1_descriptive.xlsx"),
    asTable = TRUE
    )
```


```{r}
schools |> 
  select(ends_with("count"), matches("days")) |> 
  names() |> writeLines(sep = "', '", con = "clipboard")

count_mods <- c('asthma_count', 'food_allergy_count', 'non_food_allergy_count', 
  'student_504_count', 'iep_count', 'esl_count', 'frm_count', 'homelessness_count') |>
  set_names() |> 
  map(~{
    geepack::geeglm(
      arsenal::formulize(.x, "data_schoolyear + offset(log(student_count))"),
      family = poisson(link = "log"),
      data   = schools,
      id     = school_id,
      waves  = data_schoolyear,
      corstr = "exchangeable"
      ) 
  })
count_mods |> 
  map(~{
    emmeans::emmeans(.x, ~ data_schoolyear, offset = log(100), type = "response")
    })
```


```{r}
schooldays_mod <- geepack::geeglm(
  days_per_school_present ~ data_schoolyear + offset(log(days_per_school_expected)),
      family = poisson(link = "log"),
      data   = schools |> mutate(days_per_school_present = round(days_per_school_present)),
      id     = school_id,
      waves  = data_schoolyear,
      corstr = "exchangeable"
      ) 
emmeans::emmeans(
  schooldays_mod, 
  consec ~ data_schoolyear, 
  offset = log(176), 
  type = "response"
  )  
```


```{r}
schools |> 
  select(-matches("lastyear")) |> 
  relocate(attendance, student_count, matches("days_per_sch"))
```



```{r}
'days_per_stu_expected', 'days_per_stu_present', 'days_per_stu_absent', 'days_per_school_expected', 'days_per_school_present', 'days_per_school_absent', 'days_per_stu_present_asthma', 'days_per_stu_absent_asthma', 'impl_days', 'days_per_stu_expected.lastyear', 'days_per_stu_present.lastyear', 'days_per_stu_absent.lastyear', 'days_per_school_expected.lastyear', 'days_per_school_present.lastyear', 'days_per_school_absent.lastyear', 'days_per_stu_present_asthma.lastyear', 'days_per_stu_absent_asthma.lastyear', 'impl_days.lastyear'
  
  
```




```{r}
school_change <- schools |>
  select(-matches("lastyear")) |> 
  rename_with(~ gsub("_", "", .x), where(is.numeric)) |> 
  mutate(
    cc_22.23_21.22 = case_when(
      data_schoolyear == "2022-2023" ~ 1,
      data_schoolyear == "2021-2022" ~ -1,
      .default = 0
      ),
    cc_23.24_22.23 = case_when(
      data_schoolyear == "2023-2024" ~ 1,
      data_schoolyear == "2022-2023" ~ -1,
      .default = 0
      ),
    cc_24.25_23.24 = case_when(
      data_schoolyear == "2024-2025" ~ 1,
      data_schoolyear == "2023-2024" ~ -1,
      .default = 0
      )
    ) |> 
  
  
  summarise(across(where(is.numeric), list(
    `22.23-21.22` = ~ sum(.x*cc_22.23_21.22),
    `23.24-22.23` = ~ sum(.x*cc_23.24_22.23),
    `24.25-23.24` = ~ sum(.x*cc_24.25_23.24)
    )), .by = school_id) |> 
  
  pivot_longer(where(is.numeric)) |>
  separate(name, c("variable", "years"), sep = "_") |> 
  arrange(years, variable) |> 
  mutate(years = fct_inorder(years)) |> 
  filter(variable!= "cc") |> 
  droplevels()
```

```{r}
yoy_changes <- split(school_change, school_change$variable) |> 
  map(~{
    geepack::geeglm(
      value ~ years,
      data  = .x,
      id    = school_id,
      waves = years
      )
    }) |> 
  
  map_dfr(~{
    emmeans::emmeans(.x, ~ years, infer = TRUE) |> 
      data.frame()
    }, .id = 'variable')

```

```{r}
openxlsx::read.xlsx(
  yoy_changes |> 
  separate(years, c("to", "from"), sep = "-") |> 
  mutate(
    years = glue::glue(
      "from 20{gsub('.', '-20', from, fixed = TRUE)} to 20{gsub('.', '-20', to, fixed = TRUE)}"
      ), 
    change = emmean,
    .after = 1, 
    .keep = "unused"
    ) |> 
  openxlsx::write.xlsx(
    nested_here("output", "tables", "t1_yoy_changes.xlsx"),
    asTable = TRUE
    ),
  nested_here("output", "tables", "t1_yoy_changes.xlsx")
  )
```


```{r}
yoy_changes |> 
  ggplot(aes(
    variable, emmean, group = years, 
    ymin = lower.CL, ymax = upper.CL, 
    color = factor(I(p.value <= 0.05)))
    ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_pointrange(size = 0.3) +
  # geom_linerange(position = position_dodge(width = 0.2)) +
  # facet_wrap(variable ~ ., scales = "free") +
  facet_grid(. ~ years) +
  coord_flip() +
  theme(legend.position = "bottom")
  
```












```{r}

```







```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_SCHOOLS.R"))
```

Results preview:

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_SCHOOLS.R")) |> 
  read_table() |> 
  map(head, 10) |> 
  map(knitr::kable, digits = 3) |> 
  walk(print)
```


# Utilization Detail

```{r}
tableby_with <- function(df, ...) {
  arsenal::tableby(
    ...,
    data = df
    )
}
```

```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_UTILIZATION.R"))
```

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_UTILIZATION.R")) |> 
  read_table() 
```




What predicts utilization? Poisson (or similar) for inhaler uses, 
with a student-time (in program) offset. 
Adjust for school-level chars. 

- [ ] tabulate uses per school per year

```{r}

```







- [ ] calculate student-program-time per school per year 
      (time btw program implementation and end of school year)
- [ ] calculate use rate per school per year
- [ ] plot 
    - [ ] scatter: rate over time by impl date
- [ ] model uses (count with student-time offset) 
      as predicted by school demographics,
      school characteristics, calendar time (school year), time in program

Categorize schools by rate of utilization (quartiles). Categorize 
schools by outcome rates (quartiles). Tabulate, plot by year. 


# Student Outcomes


Describe trends in student outcomes (student group level) 
over time (asthma, 504 plans, ieps, attendance).
Allow for demog-by-outcome interaction terms.


# Outcomes x Utilization


Latent class analysis/latent transition analysis. Categorize by low/high utilizers 
and low/high need. 

Describe trends in outcomes as predicted by time-in-program, utilization 
(treat as time-varying exposure?). Causal analysis w 
time-varying covariates (confounders): use IPTW for utilization?




# Write Results

```{r}
out_tables <- list(
  About = t(data.frame(
    Date   = Sys.Date(),
    Author = "Rachel Lane",
    Title  = "SOW 2 Tables"
  )) |> 
  as.data.frame() |> 
  set_names("value") |> 
  rownames_to_column("name")
  )
out_tables[["About"]] <- rbind(
  out_tables[["About"]],
  params |> 
    keep_at(~ grepl("raw", .x)) |> 
    set_names_from_child("clean_name") |> 
    map_dfr(~data.frame(
      name = "Data Sources",
      value = glue::glue("File: {.x$file}. Sheet: {coalesce(.x$sheet, '1')}. Box file ID: {.x$box_id}")
      )))

```

```{r}
for (f in list.files(
  nested_here("output", "tables"),
  "sow3"
  )) {
  d <- read_rds(nested_here("output", "tables", f))
  out_tables[[attr(d, "sheetname")]] <- d
  out_tables$About <- out_tables$About |> 
    add_case(
      name = "Contents",
      value = glue::glue('Sheet: {attr(d, "sheetname")}. Title: {attr(d, "title")}. Description: {attr(d, "description")}')
    )
  }

```

```{r}
openxlsx::write.xlsx(
  out_tables,
  nested_here("output", paste0("sow3_tables_", Sys.Date(), ".xlsx")),
  asTable = TRUE
)
```


