---
title: "CPS Data Summary: Longitudinal Outcomes"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-04-16" 
  raw_data_cps:
    value:
      file: "UIC Asthma Status Data Albuterol SoW Shell_07252025.xlsx"
      sheet: "SY22, 23, 24 and 25 Asthma Stat"
      box_id: '1938297974039'
      clean_name: "student_counts"
  raw_data_schools_metadata:
    value: 
      file: "CPS Schools CCAs Regions.xlsx"
      sheet: NULL
      box_id: '2071252778107'
      clean_name: "schools"
  raw_data_utilization:
    # value:
    #   file: "CPS 2023-2025 utilization recode.xlsx"
    #   sheet: "Sheet1"
    #   box_id: '2017350498669'
    #   clean_name: "inhaler_events"
    value: # added 2026-03-25
      file: "CPS 2023-2025 utilization recode_3-23.xlsx"
      sheet: "Utilization Data"
      box_id: '2173811730030'
      clean_name: "inhaler_events"
      # force_preproc: TRUE
  raw_data_waves: # added 2026-01-21
    value:
      file: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
      box_id: "2109640911794"
      clean_name: "waves"
  raw_sy_dates: # added 2026-03-04
    value:
      box_id: "2154041346707"
      file: "schoolyear_dates.xlsx"
      sheet: "Sheet1"
      clean_name: "school_year_dates"
      # force_preproc: true
  enrollment_count_files: # box ids
    value: 
      sy_22: "2154049626576"
      sy_23: "2154041838698"
      sy_24: "2154043506963"
      sy_25: "2154043912572"
      sy_26: "2154047037077"
  force_preproc: FALSE
knit: |
  (function(input, ...) {
      rmarkdown::render(
        input, 
        output_file   = paste0("33753_cps_data_", Sys.Date()),
        output_dir    = file.path(dirname(input), "output"),
        envir         = globalenv()
      )
  })
---

```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}
# (maybe) create subdirectories for code, data, output
walk(nested_here(c("R", "data", "output", "output/tables")), ~{
  if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
  })

# load external functions
list.files(nested_here("R"), "fx.R$", full.names = TRUE) |> 
  walk(source)
```



# Data Processing

## Process and Merge Raw Data Files

```{r, include = FALSE}
# for each raw data source...
regen_files <- params |> 
  keep_at(~ grepl("raw", .x)) |> 
  set_names_from_child("clean_name") |> 
  imap(~{
    .x |> list_assign(
      f_name         = nested_here("data", paste0("pre_", .y, ".rds")),
      force_preproc  = coalesce(
        .x |> 
          pluck("force_preproc"), 
        params$force_preproc
        )
      )
  }) |> 
  
  # if any doesn't exist or needs forced regeneration...
  keep(~ (!file.exists(.x$f_name) | .x$force_preproc)) |> 
  
  # process and write to file
  map(~{fetch_and_preproc(.x) |>  write_rds(.x$f_name)})


# make school name/id crosswalk 
# multiple source files: params$enrollment_count_files

if(params$force_preproc | 
   !file.exists(nested_here("data", "pre_school_ids.rds"))) {

  params$enrollment_count_files |> 
    imap(~{
      list(
        box_id     = .x,
        sheet      = 1,
        clean_name = gsub("sy", "school_enrollment", .y)
        )
      }) |>
    map_dfr(fetch_and_preproc) |>
    write_rds(nested_here("data", "pre_school_ids.rds"))
  
  regen_files <- rlist::list.append(
    "pre_school_ids" = list(f_name = nested_here("data", "pre_school_ids.rds"))
  )
}

# (maybe) re-merge files
if(length(regen_files) > 0) {
  source(nested_here("R", "MERGE.R"))
  }
```

```{r}

```


```{r, eval = FALSE}
# checks

# should be equivalent:
read_rds(nested_here("data", "mrg_inhaler_events.rds")) |>
    count(data_schoolyear)

read_rds(nested_here("data", "mrg_school_lvl_event_counts.rds")) |>
    summarise(n = sum(n_events), .by = data_schoolyear)
read_rds(nested_here("data", "mrg_school_lvl_student_counts.rds"))

head(read_rds(nested_here("data", "analytic_school_lvl.rds")))


# check: schools with cps data per year
read_rds(nested_here("data", "analytic_school_lvl.rds")) |>
    mutate(cps_data = !is.na(student_count)) |>
    select(data_schoolyear, cps_data) |>
    table()

# check: inhaler events by year
read_rds(nested_here("data", "analytic_school_lvl.rds")) |>
    summarise(sum(n_events), .by = data_schoolyear)

# check: inhaler events by year, with late implementers as separate category
read_rds(nested_here("data", "analytic_school_lvl.rds")) |>
    summarise(sum(n_events), .by = c(impl_schoolyear, data_schoolyear, any_events_l3)) |>
    arrange(impl_schoolyear, data_schoolyear)
```












# Describe School Chars by Year

```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_SCHOOLS.R"))
```

Results preview:

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_SCHOOLS.R")) |> 
  read_table() |> 
  map(head, 10) |> 
  map(knitr::kable, digits = 3) |> 
  walk(print)
```


# Utilization Detail

```{r}
tableby_with <- function(df, ...) {
  arsenal::tableby(
    ...,
    data = df
    )
}
```


```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_UTILIZATION.R"))
```

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_UTILIZATION.R")) |> 
  read_table() 
```







# Missing Data

```{r}
get_composition <- function(data, x, n) {
  if(is.null(n)) {
    data[["n"]] <- 1
    } else {
      data[["n"]] <- data[[n]]
    }
  
  comp <- data |> 
    summarise(n = sum(n), .by = all_of(c(x))) |> 
    na.omit() |> 
    mutate(p = n / sum(n))
  
  as.list(comp[["p"]]) |> set_names(comp[[x]])
  
  }
```


```{r}
block_impute <- function(data, var, level_probs) {
  
  if(sum(is.na(data[[var]])) == 0) {return(data)}
  
  
  data[["MISSING"]] <- is.na(data[[var]])
  data_split <- split(data, data$MISSING)
  
  miss_counts <- data_split$`TRUE` |> 
    select(-all_of(c(var)))
  
  imputed_counts <- miss_counts |> 
    mutate(
      across(matches("count_yes"), ~ .x / student_count),
      student_days_present = student_days_present/student_days_expected,
      
      imputed = map(student_count, ~{
        rmultinom(1, size = .x, prob = level_probs) |> 
          as.data.frame() |> 
          set_names("student_count") |> 
          rownames_to_column(var)
        }),
      .keep = "unused"
      ) |> 
    unnest(imputed) |> 
    filter(student_count > 0) |>
  
  mutate(
    across(matches("yes"), ~ .x*student_count),
    across(matches("yes"), ~ student_count - .x, .names = "{gsub('yes', 'no', .col)}"),
    
    student_days_expected = student_count*cps_base_days,
    
    student_days_present  = student_days_present*student_days_expected,
    student_days_absent   = student_days_expected - student_days_present
    )
  
  
  rbind(
      imputed_counts,
      data_split$`FALSE`
    ) |> 
    select(-MISSING)  
}
```

```{r}
set.seed(7453)
mod_df <- read_rds(nested_here("data", "analytic_student_lvl.rds")) |> 
  rowid_to_column() |> 
  mutate(
    student_race = fct_collapse(
      student_race,
      Other = c("Middle Eastern or North African", "Multi"))
    ) |> 
  group_nest(school_id, data_schoolyear) |> 
  mutate(
    data = map(data, function(d) {
      block_impute(
        d,
        var         = "student_race", 
        level_probs = get_composition(d, x = "student_race", n = "student_count")
        )
      }),
    data = map(data, function(d) {
      block_impute(
        d,
        var         = "student_gender", 
        level_probs = get_composition(d, x = "student_gender", n = "student_count")
        )
      })
    ) |> 
  unnest(data) |> 
  arrange(rowid)

```


```{r}
# check counts
list(
  read_rds(nested_here("data", "analytic_student_lvl.rds")),
  mod_df
  ) |> 
  map_dfr(~{
    .x |> 
      select(student_count, matches("student_days"), matches("yes$"), matches("no$")) |> 
      colSums()
    })
```






# Attendance, Asthma, IEPs, 504s

```{r}
geeify <- function(
    data, 
    sum_by     = NULL, 
    sum_across = c(
      "student_count",
      grep("student_days", names(data), value = TRUE),
      grep("504_count", names(data), value = TRUE),
      grep("iep_count", names(data), value = TRUE),
      grep("asthma_count", names(data), value = TRUE)
      )
    ) {
  
  d2 <- data |> 
    summarise(
      across(
        all_of(c(sum_across)),
      ~ round(sum(.x, na.rm = TRUE))
      ),
    .by = all_of(c(
      find_unique_by(data, c("school_id", "data_schoolyear")),
      sum_by
      ))
    ) 
  
  droplevels(d2)
}
```


```{r}
expand_grid(
  y = c(
    "cbind(student_days_present, student_days_absent)",
    "cbind(iep_count_yes, iep_count_no)",
    "cbind(student_504_count_yes, student_504_count_no)",
    "cbind(asthma_count_yes, asthma_count_no)"
  ),
  x = c(
    "1", "data_schoolyear",
    "student_race",
    "student_gender",
    "wave",
    "active_months",
    "grade_level",
    "n_events_cumulative",
    "any_events_cumulative",
    glue::glue(
      "data_schoolyear*{v}",
      v = c(
        "student_race",
        "student_gender",
        "wave",
        "active_months",
        "grade_level",
        "n_events_cumulative",
        "any_events_cumulative"
      ))
    )
  ) |> 
  mutate(
    sum_by = case_when(
      grepl("student_race", x) ~ "student_race",
      grepl("student_gender", x) ~ "student_gender"
      ),
    m_id   = 1:n(),
    f_name = nested_here("data", "models", glue::glue("m_screen_{sprintf('%04d', m_id)}.rds"))
    ) |> 
  
  write_rds(nested_here("data", "m_screen_index.rds"))
```


```{r}
read_rds(nested_here("data", "m_screen_index.rds")) |>
  filter(!file.exists(f_name)) |> 
  transpose() |>
  set_names_from_child("f_name") |> 
  
  map(safely(~{
    .x <- .x |> keep(~!is.na(.x))
    
    eval(bquote(geepack::geeglm(
      .(arsenal::formulize(.x$y, .x$x)),
      data   = geeify(mod_df, sum_by = .(pluck(.x, "sum_by"))),
      waves  = data_schoolyear,
      id     = school_id,
      corstr = "independence",
      family = binomial(link = "logit")
      )))
    
  })) |> 
  
  iwalk(write_rds)
```

```{r}
read_rds(nested_here("data", "m_screen_index.rds")) |> 
  filter(file.exists(f_name)) |> 
  pull(f_name) |> 
  map(read_rds) |> 
  transpose() |> 
  pluck("error") |> 
  compact()
```


```{r}
read_rds(nested_here("data", "m_screen_index.rds")) |> 
  filter(file.exists(f_name)) |> 
  mutate(qic = map(f_name, possibly(~{
    read_rds(.x) |> pluck("result") |> geepack::QIC() |> as.list() |> as.data.frame()
    }))) |> 
  unnest(qic) |> 
  write_rds(nested_here("data", "m_screen_qic.rds"))

read_rds(nested_here("data", "m_screen_qic.rds")) |> 
  separate(x, c("x2", "x1"), sep = "\\*", fill = "left", remove = FALSE) |>
  filter(QICu == min(QICu), .by = c(y, x1))
```

```{r}
read_rds(nested_here("data", "m_screen_index.rds")) |> 
  filter(file.exists(f_name)) |> 
  filter(!grepl("\\*", x)) |> 
  transpose() |> 
  map(possibly(~{
    m <- read_rds(.x$f_name) |> 
      pluck("result")
    list(
      emm = data.frame(
        .x, 
        emmeans::emmeans(m, .x$x, type = "response")
        ), 
      contrasts = data.frame(
        .x,
        emmeans::contrast(
          emmeans::emmeans(m, .x$x, type = "response"),
          "revpairwise"
          )
        )
    )
    })) |> 
  compact() |> 
  write_rds(nested_here("data", "m_screen_results.rds"))
```

```{r}
read_rds(nested_here("data", "m_screen_results.rds")) |> 
  map(pluck, "emm") |> 
  map_dfr(rename_if, is.factor, ~ "level") |> 
  filter(!is.na(level)) |> 
  group_split(y) |> 
  map(~{
    .x |> ggplot(aes(level, prob, ymin = lower.CL, ymax = upper.CL)) +
      geom_linerange() +
      facet_grid(x ~ ., space = "free", scales = "free") +
      coord_flip() +
      labs(title = first(.x$y))
    })
  
  
  
  map_dfr(pluck, "contrasts") |> 
  
  ggplot(aes(contrast, odds.ratio))
```





Outcome by select variables, stratified by year

 school year
  * race
  * gender
  * wave
  * active_months
  * grade_level
  * utilization

```{r}
geepack::geeglm(
  
)
```





# Longitudinal Models

* Model each outcome probability by 
  * school year
  * race
  * gender
  * wave
  * active_months
  * grade_level
  * utilization

* Check for significant interactions:
    * time by gender (504 outcome)
    * time by gender (IEP outcome)
    * time by gender (asthma outcome)
    * time by gender (attendance outcome)
    * time by race (504 outcome)
    * time by race (IEP outcome)
    * time by race (asthma outcome)
    * time by race (attendance outcome)
    



















# Write Results

```{r}
out_tables <- list(
  About = t(data.frame(
    Date   = Sys.Date(),
    Author = "Rachel Lane",
    Title  = "SOW 2 Tables"
  )) |> 
  as.data.frame() |> 
  set_names("value") |> 
  rownames_to_column("name")
  )
out_tables[["About"]] <- rbind(
  out_tables[["About"]],
  params |> 
    keep_at(~ grepl("raw", .x)) |> 
    set_names_from_child("clean_name") |> 
    map_dfr(~data.frame(
      name = "Data Sources",
      value = glue::glue("File: {.x$file}. Sheet: {coalesce(.x$sheet, '1')}. Box file ID: {.x$box_id}")
      )))

```

```{r}
for (f in list.files(
  nested_here("output", "tables"),
  "sow3"
  )) {
  d <- read_rds(nested_here("output", "tables", f))
  out_tables[[attr(d, "sheetname")]] <- d
  out_tables$About <- out_tables$About |> 
    add_case(
      name = "Contents",
      value = glue::glue('Sheet: {attr(d, "sheetname")}. Title: {attr(d, "title")}. Description: {attr(d, "description")}')
    )
  }

```

```{r}
openxlsx::write.xlsx(
  out_tables,
  nested_here("output", paste0("sow3_tables_", Sys.Date(), ".xlsx")),
  asTable = TRUE
)
```


