---
title: "CPS Data Summary"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-04-06" 
  raw_data_cps:
    value:
      file: "UIC Asthma Status Data Albuterol SoW Shell_07252025.xlsx"
      sheet: "SY22, 23, 24 and 25 Asthma Stat"
      box_id: '1938297974039'
      clean_name: "student_counts"
  raw_data_schools_metadata:
    value: 
      file: "CPS Schools CCAs Regions.xlsx"
      sheet: NULL
      box_id: '2071252778107'
      clean_name: "schools"
  raw_data_utilization:
    # value:
    #   file: "CPS 2023-2025 utilization recode.xlsx"
    #   sheet: "Sheet1"
    #   box_id: '2017350498669'
    #   clean_name: "inhaler_events"
    value: # added 2026-03-25
      file: "CPS 2023-2025 utilization recode_3-23.xlsx"
      sheet: "Utilization Data"
      box_id: '2173811730030'
      clean_name: "inhaler_events"
      force_preproc: TRUE
  raw_data_waves: # added 2026-01-21
    value:
      file: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
      box_id: "2109640911794"
      clean_name: "waves"
  raw_sy_dates: # added 2026-03-04
    value:
      box_id: "2154041346707"
      file: "schoolyear_dates.xlsx"
      sheet: "Sheet1"
      clean_name: "school_year_dates"
      # force_preproc: true
  enrollment_count_files: # box ids
    value: 
      sy_22: "2154049626576"
      sy_23: "2154041838698"
      sy_24: "2154043506963"
      sy_25: "2154043912572"
      sy_26: "2154047037077"
  force_preproc: FALSE
knit: |
  (function(input, ...) {
      rmarkdown::render(
        input, 
        output_file   = paste0("33753_cps_data_", Sys.Date()),
        output_dir    = file.path(dirname(input), "output"),
        envir         = globalenv()
      )
  })
---

```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}
# (maybe) create subdirectories for code, data, output
walk(nested_here(c("R", "data", "output", "output/tables")), ~{
  if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
  })

# load functions
list.files(nested_here("R"), "fx.R$", full.names = TRUE) |> 
  walk(source)
```



# Data Processing

## Process and Merge Raw Data Files

```{r, include = FALSE}
# for each raw data source...
regen_files <- params |> 
  keep_at(~ grepl("raw", .x)) |> 
  set_names_from_child("clean_name") |> 
  imap(~{
    .x |> list_assign(
      f_name         = nested_here("data", paste0("pre_", .y, ".rds")),
      force_preproc  = coalesce(
        .x |> 
          pluck("force_preproc"), 
        params$force_preproc
        )
      )
  }) |> 
  
  # if any doesn't exist or needs forced regeneration...
  keep(~ (!file.exists(.x$f_name) | .x$force_preproc)) |> 
  
  # process and write to file
  map(~{fetch_and_preproc(.x) |>  write_rds(.x$f_name)})

# (maybe) re-merge files
if(length(regen_files) > 0) {
  source(nested_here("R", "MERGE.R"))
  }
```

# Describe School Chars by Year

```{r}
write_table <- function(x, loc = nested_here("output", "tables")) {
  n_extant <- length(list.files(loc, "^t.*rds$"))
  write_rds(
    x,
    file.path(loc, glue::glue("t{sprintf('%03d', n_extant+1)}.rds"))
  )
}
```

```{r}
# schools overall, annual data
school_tby(
  y_var = "",
  df    = read_rds(nested_here("data", "analytic_school_lvl.rds")),
  ) |> 
  structure(
    title       = "School Characteristics by School Year",
    sheetname   = "descr_school_char_by_year",
    description = "Summary statistics of school-level characteristics and demographics. Variables from CPS (e.g., % with asthma) are tabulated separately for each school year (2021-2022 to 2024-2025)."
    ) |>
  
  write_rds(nested_here("output", "tables", "sow3_t01a_chars_by_sy.rds"))
```


```{r}
# test changes from year to year (all pairwise)
read_rds(nested_here("data", "analytic_school_lvl.rds")) |> 
  filter(!is.na(student_count)) |> 
  droplevels() |>
  pivot_longer(
    c(student_count, average_attendance, matches("pct")),
    names_to = "x"
    ) |> 

  pivot_wider(
    names_from = data_schoolyear,
    id_cols    = c(school_id, x)
    ) |> 
  
  mutate(
   `(2022-2023) - (2021-2022)` = `2022-2023` - `2021-2022`,
   `(2023-2024) - (2021-2022)` = `2023-2024` - `2021-2022`,
   `(2023-2024) - (2022-2023)` = `2023-2024` - `2022-2023`,
   `(2024-2025) - (2021-2022)` = `2024-2025` - `2021-2022`,
   `(2024-2025) - (2022-2023)` = `2024-2025` - `2022-2023`,
   `(2024-2025) - (2023-2024)` = `2024-2025` - `2023-2024`
   ) |> 
  
  select(school_id, x, matches("\\(")) |> 
  pivot_longer(where(is.numeric), names_to = "contrast", values_to = "abs_change") |> 
  
  summarise(
    n_schools = sum(!is.na(abs_change)),
    broom::tidy(wilcox.test(abs_change, exact = TRUE, conf.int = TRUE)),
    .by       = c(x, contrast)
  ) |> 
  rename(med.change = estimate, med.change.conf.low = conf.low, med.change.conf.high = conf.high) |> 
  mutate(
    p.adjust        = p.adjust(p.value, "holm"), 
    p.adjust.n      = sum(!is.na(p.value)),
    p.adjust.method = "holm",
    .after          = p.value,
    .by             = x
    )  |> 
  
  structure(
    title       = "Test Change in School Characteristics Over Time",
    sheetname   = "test_school_char_change",
    description = "Annual school-level numeric variables (e.g., attendance, % with asthma) were tested for evidence of change over time. For each variable, within-school change was calculated from schoolyear to schoolyear (all pairwise schoolyear comparisons). School-level change between each pair of years was then tested for symmetry around zero (i.e., no change over time) by a one-sample Wilcoxon signed-rank test. p-values for the set of 6 pairwise year comparisons for each variable were adjusted by the Holm method."
    ) |>
  
    write_rds(nested_here("output", "tables", "sow3_t01b_test_chars_by_sy.rds"))
```







# Describe School Chars (2023-2024) by Inhaler Events

```{r}
# any events, 2023-2024 only, separating late implementers
school_tby(
  y_var = "any_events_l3",
  df    = read_rds(nested_here("data", "analytic_school_lvl.rds")) |> 
    filter(data_schoolyear %in% c("2023-2024")),
   tby_ctrl = arsenal::tableby.control(
      test           = FALSE,
      digits.n	      = NA,
      numeric.stats  = c("medianrange", "q1q3",  "meansd")
    ),
  ) |> 
  
  structure(
    title       = "School Characteristics by Leven of Inhaler Utilization, SY 2023-2024",
    sheetname   = "descr_util_groups_23_24",
    description = "Summary statistics of school-level characteristics and demographics in the 2023-2024 schoolyear. Schools are grouped by inhaler utilization level in the 2023-2024 school year (0 events vs. 1+ events vs. N/A).") |> 
  
    write_rds(nested_here("output", "tables", "sow3_t02a_chars_by_util.rds"))
```


```{r}
full_join(
  
  c("grade_level", "region_3grp",  "region_2grp" ) |> 
    set_names() |> 
    map_dfr(~{
      test_categ(.x, groups = "any_events_l3", data = read_df_sy("2023-2024"))
    }, .id = "x"),
  
  c("student_count", "average_attendance", 
    grep("pct", names(read_df_sy("2023-2024")), value = TRUE)) |> 
    set_names() |>  
    map_dfr(~{
      test_ordered(.x, groups = "any_events_l3", data = read_df_sy("2023-2024"))
    }, .id = "x")
  
) |> 
  mutate(data_schoolyear = "2023-2024", .before = 1) |> 
  relocate(data_schoolyear, x, x_pair, group1, group2, n, p.value, matches("p.adj")) |> 
  
    structure(
    title       = "Test Differences in School Characteristics by Level of Inhaler Utilization, SY 2023-2024",
    sheetname   = "test_util_group_diffs_23_24",
    description = "Three-group and pairwise comparisons of school characteristics by inhaler utilization level in the SY 2023-2024 schoolyear. Numeric variables are compared across groups (0 events, 1+ events, NA) by Kruskal-Wallis tests (all 3 groups) and pairwise Wilcoxon rank-sum tests (each pair of 2 groups). Categorical variables (e.g., grade level) are compared across groups (0 events, 1+ events, NA) Fisher's exact tests (all 3 groups and each pair of 2 groups). p-values for the set of three pairwise comparisons for each variable were adjusted by the Holm method.") |> 
  
  write_rds(nested_here("output", "tables", "sow3_t02b_test_chars_by_util.rds"))
```




# Utilization Detail

```{r}

```


```{r}
full_join(
    read_rds(nested_here("data", "mrg_school_chars.rds")),
    read_rds(nested_here("data", "mrg_inhaler_events.rds")),
    by = "school_id"
    ) |> 
    select(-where(is.character)) |> 
    select(-matches(" ")) |> 
    select(-matches("\\.\\.\\.")) |> 
    relocate(-any_of(c("impl_date", "school_id", "admin_protocol", "rcp_role",
    "rcp_role_2", "hypo_disp_rcd", "hosp", "transport", matches("missing"),
    "used_before", "...88", "data_season", "source"
    ))) |> 
  
  (\(.) 
   
   

arsenal::tableby(
  arsenal::formulize(
    "data_schoolyear", c(
    
    # recipient
    "rcp_sex",  "rcp_race",  "rcp_eth", # "rcp_age",
    "rcp_role",  "sym_new",  "rcp_dx_pre",  "inh_use_prev",

    # incident
    "season", "why_given",
    
    # school chars
      "wave", "grade_level",
    
    # disposition
    # hypo_disp_rcd + true_disp_rcd,
  
      names(.))
    ),
  # test = FALSE,
  data = .,
  control = arsenal::tableby.control(
    test           = FALSE,
    digits.n	      = NA,
    numeric.stats  = c("medianrange", "q1q3",  "meansd")
    )
  # total = FALSE
  ) |>
  arsenal:::summary.tableby(text = TRUE, labelTranslations = list(rcp_age = "Age")) |>
  as.data.frame()

)()
  # write_output("Table2_Utilization")
```


```{r}

```









## Attendance, Asthma, IEPs, 504s


# Longitudinal Models

* Check for significant interactions:
    * time by gender (504 outcome)
    * time by gender (IEP outcome)
    * time by gender (asthma outcome)
    * time by gender (attendance outcome)
    * time by race (504 outcome)
    * time by race (IEP outcome)
    * time by race (asthma outcome)
    * time by race (attendance outcome)
    
    Model each outcome probability by school year, race, gender, 
active_months, wave, grade_level, utilization







# Write Results

```{r}
out_tables <- list(
  About = t(data.frame(
    Date   = Sys.Date(),
    Author = "Rachel Lane",
    Title  = "SOW 2 Tables"
  )) |> 
  as.data.frame() |> 
  set_names("value") |> 
  rownames_to_column("name")
  )
out_tables[["About"]] <- rbind(
  out_tables[["About"]],
  params |> 
    keep_at(~ grepl("raw", .x)) |> 
    set_names_from_child("clean_name") |> 
    map_dfr(~data.frame(
      name = "Data Sources",
      value = glue::glue("File: {.x$file}. Sheet: {coalesce(.x$sheet, '1')}. Box file ID: {.x$box_id}")
      )))

```

```{r}
for (f in list.files(
  nested_here("output", "tables"),
  "sow3"
  )) {
  d <- read_rds(nested_here("output", "tables", f))
  out_tables[[attr(d, "sheetname")]] <- d
  out_tables$About <- out_tables$About |> 
    add_case(
      name = "Contents",
      value = glue::glue('Sheet: {attr(d, "sheetname")}. Title: {attr(d, "title")}. Description: {attr(d, "description")}')
    )
  }

```

```{r}
openxlsx::write.xlsx(
  out_tables,
  nested_here("output", paste0("sow3_tables_", Sys.Date(), ".xlsx")),
  asTable = TRUE
)
```


