---
title: "CPS Data Summary: Longitudinal Outcomes"
author: "Rachel Lane"
date: "`r format(Sys.Date(), format = '%B %d, %Y')`"
params:
  parent_dir: "a12_cps_data" # relative to project root
  version: "2026-09-23"
  analytic_data_dir_id: "403977851006"
  preproc: false
  raw_data_cps_counts:
    value:
      box_id: "2482165585966"
      filename: "UIC Asthma Status Data Albuterol SoW Shell_09172026.xlsx"
      sheet: "SY22, 23, 24, 25 and 26 Asthma "
  raw_data_utilization:
    value: 
      box_id: '2173811730030'
      filename: "CPS 2023-2025 utilization recode_3-23.xlsx"
      sheet: "Utilization Data"
  raw_data_waves:
    value:
      box_id: "2109640911794"
      filename: "CPS Ramp Up School List_MostUpdated.xlsx"
      sheet: "Sheet1"
  raw_data_eji:
    value:
      box_id: "2237279411094"
      filename: "CPS_EJI_Excel.xls"
      sheet: "AllCPS2122_Merge"
# knit: |
#   (function(input, ...) {
#       rmarkdown::render(
#         input, 
#         output_file   = paste0("33753_cps_data_", Sys.Date()),
#         output_dir    = file.path(dirname(input), "output"),
#         envir         = globalenv()
#       )
#   })
---

```{r K, echo = FALSE, include = FALSE}
# define document options
knitr::opts_chunk$set(
  echo    = FALSE, eval    = TRUE, 
  message = FALSE, warning = FALSE
  )

params <- params
```

```{r, include = FALSE}
# load libraries
library(tidyverse)
boxr::box_auth(cache = here::here(".boxr-oauth"))
```

```{r chunk_opts, echo = FALSE, include = FALSE}
# define formatting options
options(knitr.kable.NA = '')
thm <- theme_set(theme_bw())
```

```{r}
# function for easier navigation of relative refs
nested_here <- function(
    ...,    
    parent_dir = params$parent_dir, 
    version    = params$version
    ) {
    here::here(parent_dir, version, ...)
}
```

```{r}
# # (maybe) create subdirectories for code, data, output
walk(nested_here(c("R", "data", "output", "output/tables")), ~{
  if(!dir.exists(.x)) {dir.create(.x, recursive = TRUE)}
  })
```

```{r}
# define custom functions
nested_here("R", c("box_fx.R")) |> 
  walk(source)
```


```{r}
# set defaults for box functions
fx_defaults <- list(
  box_file_attr     = default::default(box_file_attr),
  box_write_if_diff = default::default(box_write_if_diff)
  )
default::default(box_file_attr) <- list(
  dir_id   = params$analytic_data_dir_id
  )
default::default(box_write_if_diff) <- list(
  log_file = "data_log.csv",
  dir_id   = params$analytic_data_dir_id,
  force    = FALSE
  )
```


```{r, eval = FALSE, include = FALSE}
# # fetch external/public data
# source(nested_here("R", "r01a_FETCH_API_DATA.R"))
# # 
# # # reformat data extract from cps; tabulate at school and schoolyear levels
# source(nested_here("R", "r01b_PREPROC_CPS_FILE.R"))
# # 
# # # prepare wave assignments
# source(nested_here("R", "r01c_PREPROC_WAVES.R"))
# # 
# # # prepare EJI RPL SER variable
# source(nested_here("R", "r01d_PREPROC_EJI.R"))
# # 
# # # count inhaler incidents
# source(nested_here("R", "r01e_PREPROC_UTILIZATION.R"))
# 
# # combine data sources into analytic dataset
# source(nested_here("R", "r02_MERGE_DATA_SOURCES.R"))
```



# Describe School Chars by Year

```{r}
schools <- read_rds(nested_here("data/cps_schools_analytic.rds")) 
```


```{r, include = FALSE}
# library(arsenal)
source(nested_here("R", "r04a_DESCRIBE_SCHOOLS.R"))
source(nested_here("R", "r04b_DESCRIBE_WAVES.R"))
source(nested_here("R", "r05_ANNUAL_CHANGE.R"))
# source(nested_here("R", "r10_IPTW_NB.R"))
```


```{r}
list(
  "T1a. Schools"           = read_rds(nested_here("output", "describe_schools.rds")),
  "T1b. Waves"             = read_rds(nested_here("output", "describe_waves.rds")),
  "T2. Annual Rate Change" = read_rds(nested_here("output", "annual_change.rds"))
  ) |> 
  
  openxlsx::write.xlsx(
    nested_here("output", "tables", glue::glue("{Sys.Date()}_t1_descriptive_and_yoy.xlsx")),
    asTable = TRUE
  )
```








```{r}


```







```{r}
# read_rds(nested_here("output", "t1_descriptive.rds")) |> 
#   openxlsx::write.xlsx(
#     nested_here("output", "tables", "t1_descriptive.xlsx"),
#     asTable = TRUE
#     )
```



```{r}
read_rds(nested_here("output", "yoy_unadj_rates.Rds")) |> 
  
  filter(variable == "asthma_count") |> 

```




```{r}
enframe(unlist(yoy_change$offset), name = "variable", value = "rate_offset"),
  by = "variable"
  ) |> 
  relocate(rate_offset, .after = variable) |> 
  select(-any_of(c("null", "t.ratio", "p.value")))
  
```



```{r}
openxlsx::read.xlsx(
  yoy_changes |> 
  separate(years, c("to", "from"), sep = "-") |> 
  mutate(
    years = glue::glue(
      "from 20{gsub('.', '-20', from, fixed = TRUE)} to 20{gsub('.', '-20', to, fixed = TRUE)}"
      ), 
    change = emmean,
    .after = 1, 
    .keep = "unused"
    ) |> 
  openxlsx::write.xlsx(
    nested_here("output", "tables", "t1_yoy_changes.xlsx"),
    asTable = TRUE
    ),
  nested_here("output", "tables", "t1_yoy_changes.xlsx")
  )
```


```{r}
yoy_changes |> 
  ggplot(aes(
    variable, emmean, group = years, 
    ymin = lower.CL, ymax = upper.CL, 
    color = factor(I(p.value <= 0.05)))
    ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  geom_pointrange(size = 0.3) +
  # geom_linerange(position = position_dodge(width = 0.2)) +
  # facet_wrap(variable ~ ., scales = "free") +
  facet_grid(. ~ years) +
  coord_flip() +
  theme(legend.position = "bottom")
  
```












```{r}

```







```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_SCHOOLS.R"))
```

Results preview:

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_SCHOOLS.R")) |> 
  read_table() |> 
  map(head, 10) |> 
  map(knitr::kable, digits = 3) |> 
  walk(print)
```


# Utilization Detail

```{r}
tableby_with <- function(df, ...) {
  arsenal::tableby(
    ...,
    data = df
    )
}
```

```{r, eval = FALSE}
source(nested_here("R", "DESCRIBE_UTILIZATION.R"))
```

```{r}
find_written_fnames(nested_here("R", "DESCRIBE_UTILIZATION.R")) |> 
  read_table() 
```




What predicts utilization? Poisson (or similar) for inhaler uses, 
with a student-time (in program) offset. 
Adjust for school-level chars. 

- [ ] tabulate uses per school per year

```{r}

```







- [ ] calculate student-program-time per school per year 
      (time btw program implementation and end of school year)
- [ ] calculate use rate per school per year
- [ ] plot 
    - [ ] scatter: rate over time by impl date
- [ ] model uses (count with student-time offset) 
      as predicted by school demographics,
      school characteristics, calendar time (school year), time in program

Categorize schools by rate of utilization (quartiles). Categorize 
schools by outcome rates (quartiles). Tabulate, plot by year. 


# Student Outcomes


Describe trends in student outcomes (student group level) 
over time (asthma, 504 plans, ieps, attendance).
Allow for demog-by-outcome interaction terms.


# Outcomes x Utilization


Latent class analysis/latent transition analysis. Categorize by low/high utilizers 
and low/high need. 

Describe trends in outcomes as predicted by time-in-program, utilization 
(treat as time-varying exposure?). Causal analysis w 
time-varying covariates (confounders): use IPTW for utilization?




# Write Results

```{r}
out_tables <- list(
  About = t(data.frame(
    Date   = Sys.Date(),
    Author = "Rachel Lane",
    Title  = "SOW 2 Tables"
  )) |> 
  as.data.frame() |> 
  set_names("value") |> 
  rownames_to_column("name")
  )
out_tables[["About"]] <- rbind(
  out_tables[["About"]],
  params |> 
    keep_at(~ grepl("raw", .x)) |> 
    set_names_from_child("clean_name") |> 
    map_dfr(~data.frame(
      name = "Data Sources",
      value = glue::glue("File: {.x$file}. Sheet: {coalesce(.x$sheet, '1')}. Box file ID: {.x$box_id}")
      )))

```

```{r}
for (f in list.files(
  nested_here("output", "tables"),
  "sow3"
  )) {
  d <- read_rds(nested_here("output", "tables", f))
  out_tables[[attr(d, "sheetname")]] <- d
  out_tables$About <- out_tables$About |> 
    add_case(
      name = "Contents",
      value = glue::glue('Sheet: {attr(d, "sheetname")}. Title: {attr(d, "title")}. Description: {attr(d, "description")}')
    )
  }

```

```{r}
openxlsx::write.xlsx(
  out_tables,
  nested_here("output", paste0("sow3_tables_", Sys.Date(), ".xlsx")),
  asTable = TRUE
)
```


